---
title: "iNaturalist Observations in GGPlot2"
author: "Wendy Anthony"
date: "`r file.info(knitr::current_input())$ctime`"
output:
  html_document:
    highlight: zenburn
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
# output will scroll if higher than 300px
# only if df has few columns to all show in output box
.scroll-100 {
  max-height: 300px; # for vertical scroll
  max-width: 1500px; # for horizontal scroll
  overflow-y: auto; # for vertical scroll
  overflow-x: auto; # for horizontal scroll  
  background-color: inherit;
}
```

# Load Package Libraries
```{r load-package}
library(lubridate) # date format
library(stringr) # separate datetime column
library(ggplot2)
library(ggiraph) # for tool tip interactivity
library(tidyverse)
```

# Esquisse to setup ggplot 
  https://github.com/dreamRs/esquisse

# Import data from RDS file
created from /Users/wanthony/Documents/R/iNaturalist-shiny-leaflet-app/app.R
```{r, class.output="scroll-100"}

wa_inat <- readRDS("wa_inat_userstats_research_append.RDS")
```

# Data
## Column Names
```{r, class.output="scroll-100"}
colnames(wa_inat)
```

# Date formatting with lubridate
## class of ```observed_on_string```
```{r, class.output="scroll-100"}
class(wa_inat$observed_on_string) # character
```

# Date formatting with lubridate
## class of ```datetime```
```{r, class.output="scroll-100"}
class(wa_inat$datetime) # character
```

# Date formatting with lubridate
## str
```{r, class.output="scroll-100"}
str(wa_inat)
```

## create new columns of empty cells
```{r, class.output="scroll-100"}
wa_inat["date"] <- ""
wa_inat["date_time"] <- ""
wa_inat["time"] <- ""
```

## As Date ```datetime```
```{r, class.output="scroll-100"}
wa_inat$date <- as_date(wa_inat$datetime)
str(wa_inat)
```

# separate datetime column using ```stringr```
## Split date time into 3 separate columns
```{r, class.output="scroll-100"}
wa_inat[c('date', 'time', 'UTC')] <- str_split_fixed(wa_inat$datetime, ' ', 3)
```

### concatenate 2 columns
```{r, class.output="scroll-100"}
wa_inat$date_time <- paste(wa_inat$date, wa_inat$time)
```

# convert to date format ??? 
```{r, class.output="scroll-100"}
wa_inat$date_time <- as_datetime(wa_inat$date_time)  # this time is wrong
```

# Dataframe info
## Class
```{r, class.output="scroll-100"}
class(wa_inat$date) # date
class(wa_inat$date_time) # POSIXct
```

## Structure of db
```{r, class.output="scroll-100"}
str(wa_inat)
```

## Column names
```{r, class.output="scroll-100"}
colnames(wa_inat)
```

# Subset by choosing columns to include
```{r, class.output="scroll-100"}
wa_inat_sel <- subset(wa_inat, select = c(1,4,5,6,8,9,10,11,15,18,40,41,43,44,45))
```

# clean up strings
* place_guess has too many characters, including province, etc  
* difficult to use as plot axis labels
https://statisticsglobe.com/extract-substring-before-or-after-pattern-in-r

## Subset pattern to string content before removed text
# create new column to use to hold strings after pattern removal
```{r, class.output="scroll-100"}
wa_inat_sel$place <- sub(", BC, Canada", "", wa_inat_sel$place_guess)
```

## in new column subset to remove excess place name
```{r, class.output="scroll-100"}
wa_inat_sel$place <- sub(", BC", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Oak Bay", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", View Royal", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", James Bay, Victoria", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Capital", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Colwood", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Victoria", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Metchosin", "", wa_inat_sel$place)
wa_inat_sel$place <- sub("James Bay", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", CA", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Saanich", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Beaver Lake Regional Park", "", wa_inat_sel$place)
wa_inat_sel$place <- sub("-BC, CA", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", British Columbia, Canada", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Juan de Fuca BC Parks", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(" Juan de Fuca Provincial Park", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(" Juan de Fuca BC Parks", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(",  Strathcona Provincial Park", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Strathcona Provincial Park", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(" Juan de Fuca Trail BC Parks", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(" V9A, Canada", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", West Coast Road", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Strathcona-Westmin Provincial Park", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", North Saanich", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Jordan River Regional Park", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Gowlland Tod Provincial Park, Brentwood Bay", "", wa_inat_sel$place)
wa_inat_sel$place <- sub(", Sooke", "", wa_inat_sel$place)
```

# find unique values in columns
```{r, class.output="scroll-100"}
unique_place <- unique(wa_inat_sel$place)
class(unique_place) # character
unique_place_df <- data.frame(unique_place)
class(unique_place_df) # data.frame
```

## sort unique places in alphabetical order
```{r, class.output="scroll-100"}
unique_place_df_order <- unique_place_df$unique_place[order(unique_place_df$unique_place)]
```

## create dataframe from unique placenames
```{r, class.output="scroll-100"}
class(unique_place_df_order) # character
unique_place_df_order_df <- data.frame(unique_place_df_order)
```

## replace column values in wa_inat_sel$place to ensure only unique values exist for each place
```{r, class.output="scroll-100"}
wa_inat_sel["place"][wa_inat_sel["place"] == "Albert Head Lagoon Regional Park"] <- "Albert Head Lagoon"
wa_inat_sel["place"][wa_inat_sel["place"] == "Beacon Hill Park "] <- "Beacon Hill Park"
wa_inat_sel["place"][wa_inat_sel["place"] == "Beaver Lake Park Ponds"] <- "Beaver Lake Ponds"
wa_inat_sel["place"][wa_inat_sel["place"] == "Canada"] <- "University of Victoria"
wa_inat_sel["place"][wa_inat_sel["place"] == "Cattle Point Park"] <- "Cattle Point"
wa_inat_sel["place"][wa_inat_sel["place"] == "Cattle Point, Uplands Park"] <- "Cattle Point"
wa_inat_sel["place"][wa_inat_sel["place"] == "Cervus Creek Strathcona Provincial Park"] <- "Cervus Creek"
wa_inat_sel["place"][wa_inat_sel["place"] == "China Beach"] <- "China Beach"
wa_inat_sel["place"][wa_inat_sel["place"] == "China Beach,"] <- "China Beach"
wa_inat_sel["place"][wa_inat_sel["place"] == "Cole's Bay Regional Park"] <- "Coles Bay Regional Park"
wa_inat_sel["place"][wa_inat_sel["place"] == "Fort Rodd Hill Park Colwood"] <- "Fort Rodd Hill Park"
wa_inat_sel["place"][wa_inat_sel["place"] == "Esquimalt Lagoon, Royal Roads"] <- "Royal Roads"
wa_inat_sel["place"][wa_inat_sel["place"] == "Royal Roads Forest"] <- "Royal Roads"
wa_inat_sel["place"][wa_inat_sel["place"] == "royal_roads"] <- "Royal Roads"
wa_inat_sel["place"][wa_inat_sel["place"] == "Goldstream Park"] <- "Goldstream Provincial Park"
wa_inat_sel["place"][wa_inat_sel["place"] == "Gorge -Tillicum Area"] <- "Gorge Tillicum Area"
wa_inat_sel["place"][wa_inat_sel["place"] == "Harling PointOak Bay"] <- "Harling Point"
wa_inat_sel["place"][wa_inat_sel["place"] == "Island View Beach Regional Park"] <- "Island View Beach"
wa_inat_sel["place"][wa_inat_sel["place"] == "Juan de Fuca Trail BC Parks"] <- "Juan de Fuca Trail"
wa_inat_sel["place"][wa_inat_sel["place"] == "McKenzie Bight Gowlland Tod Provincial Park"] <- "McKenzie Bight"
wa_inat_sel["place"][wa_inat_sel["place"] == "McKenzie Bight, Gowlland Tod Provincial Park"] <- "McKenzie Bight"
wa_inat_sel["place"][wa_inat_sel["place"] == "Mt Tolmie, VIctoria"] <- "Mt Tolmie"
wa_inat_sel["place"][wa_inat_sel["place"] == "Pacific Rim National Park Reserve, Alberni-Clayoquot"] <- "Pacific Rim National Park"
wa_inat_sel["place"][wa_inat_sel["place"] == "Pacific Rim First Nations Trail"] <- "Pacific Rim National Park Reserve"
wa_inat_sel["place"][wa_inat_sel["place"] == "Pacific Rim National Park Reserve"] <- "Pacific Rim National Park"
wa_inat_sel["place"][wa_inat_sel["place"] == "Partridge Hills Gowlland Tod Provincial Park"] <- "Partridge Hills"
wa_inat_sel["place"][wa_inat_sel["place"] == "Rithet's Bog Park"] <- "Rithets Bog Park"
wa_inat_sel["place"][wa_inat_sel["place"] == "Thetis Lake Regional Park"] <- "Thetis Lake"
wa_inat_sel["place"][wa_inat_sel["place"] == "Thetis Lake Park"] <- "Thetis Lake"
wa_inat_sel["place"][wa_inat_sel["place"] == "48.464627, -123.315407"] <- "University of Victoria"
wa_inat_sel["place"][wa_inat_sel["place"] == "48.46549166666667, -123.30282833333332"] <- "University of Victoria"
wa_inat_sel["place"][wa_inat_sel["place"] == ""] <- "University of Victoria"
wa_inat_sel["place"][wa_inat_sel["place"] == ""] <- "University of Victoria"
```

## recheck unique vales
```{r, class.output="scroll-100"}
unique_place_after <- unique(wa_inat_sel$place)
unique_place_after_df <- data.frame(unique_place_after)
```

## re sort new unique values
```{r, class.output="scroll-100"}
unique_place_after_df_order <- unique_place_after_df$unique_place[order(unique_place_after_df$unique_place)]
```

### create dataframe from new unique place names
```{r, class.output="scroll-100"}
class(unique_place_after_df_order) # character
unique_place_after_df_order_df <- data.frame(unique_place_after_df_order)
class(unique_place_after_df_order_df) # dataframe
```

# GGplot2 plots
## ggplot scientific name and place
```{r, class.output="scroll-100"}
ggplot(wa_inat_sel) +
  aes(x = scientific_name, y = place) +
  geom_tile(size = 1.2) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  scale_y_discrete(limits=rev) + # reverses order of labels
  labs(title = "My iNaturalist Research Grade Observations") +
  # theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 4), axis.text.y = element_text(size = 4))
```

### save plot image
```{r, class.output="scroll-100"}
getwd()
ggsave("wa_inat_sel_scinam.jpg")
```

## ggplot Genus and place
```{r, class.output="scroll-100"}
ggplot(wa_inat_sel) +
  aes(x = Genus, y = place) +
  geom_tile(size = 1.2) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  scale_y_discrete(limits=rev) + # reverses order of labels
  labs(title = "My iNaturalist Research Grade Observations") +
  # theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 4),
        axis.text.y = element_text(size = 4))
```

### save plot image
```{r, class.output="scroll-100"}
ggsave("wa_inat_sel_genus_dodge.jpg")
```


## ggplot Genus and place with flipped coordinates
```{r, class.output="scroll-100"}
ggplot(wa_inat_sel) +
  aes(x = Genus, y = place) +
#   aes(x = Genus, y = place, fill = Genus) +   # ???????????
  geom_tile(size = 1.2) +
#  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  scale_x_discrete(limits=rev) + # reverses order of labels
  labs(title = "My iNaturalist Research Grade Observations") +
  # theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 4), 
        axis.text.y = element_text(angle = 45, vjust = 1, hjust = 1, size = 4)) +
  coord_flip()
```


### save plot image
```{r, class.output="scroll-100"}
ggsave("wa_inat_sel_genus.jpg")

```

## ggplot Genus and place with flipped coordinates and fill ????
```{r, class.output="scroll-100"}
ggplot(wa_inat_sel) +
aes(x = Genus, y = place, fill = Genus) +   # ???????????
  geom_tile(size = 1.2) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  labs(title = "My iNaturalist Research Grade Observations") +
  # theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 4), 
        axis.text.y = element_text(angle = 45, vjust = 1, hjust = 1, size = 4)) +
  coord_flip()
```

### save plot image
```{r, class.output="scroll-100"}
ggsave("wa_inat_sel_genus_fill.jpg")  # ???????
```


# Tool tip interactivity using ```ggiraph```
* TOOL TIP STILL NOT WORKING in plot ????
## make tooltip column
https://www.infoworld.com/article/3626911/easy-interactive-ggplot-graphs-in-r-with-ggiraph.html
```{r, class.output="scroll-100"}
wa_inat_sel <- wa_inat_sel %>%
  mutate(
    tooltip_text = paste0(toupper(place),  ": ", "\n",
                          Genus, ", ", "\n", "Date: ", observed_on_string)
  )
```

## ggplot with tooltip, not working ???

```{r, class.output="scroll-100"}
wa_inat_sel_tooltip <- ggplot(wa_inat_sel) +
  aes(x = Genus, y = place,
      tooltip = tooltip_text, data_id = Genus) +
  geom_tile(size = 1.2) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  scale_y_discrete(limits=rev) + # reverses order of labels
  labs(title = "My iNaturalist Research Grade Observations") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 4), axis.text.y = element_text(size = 4))
wa_inat_sel_tooltip
```

### save as svg ggobj with tooltip, though not working ???
```{r, class.output="scroll-100"}
wa_inat_sel_genus_tootip.svg <- girafe(ggobj = wa_inat_sel_tooltip, width_svg = 5, height_svg = 4)
```

### save as html page with ```htmlwidgets``` 
* ... though tool tip still not showing
```{r, class.output="scroll-100"}
htmlwidgets::saveWidget(wa_inat_sel_genus_tootip.svg, "wa_inat_sel_genus_tootip.svg.html",
                        selfcontained = TRUE)
```

# Create new column with place names in lower case
https://www.datasciencemadesimple.com/convert-to-lower-case-in-r-dataframe-column-2-2/
```{r, class.output="scroll-100"}
wa_inat_sel$place_lower <-tolower(wa_inat_sel$place)
```

## replace space with underscore
* do multiple times for places with more than one blank space
```{r, class.output="scroll-100"}
wa_inat_sel$place_lower <- sub(" ", "_", wa_inat_sel$place_lower)
wa_inat_sel$place_lower <- sub(" ", "_", wa_inat_sel$place_lower)
wa_inat_sel$place_lower <- sub(" ", "_", wa_inat_sel$place_lower)
```


### recheck unique vales
```{r, class.output="scroll-100"}
place_lower <- unique(wa_inat_sel$place_lower)
place_lower_df <- data.frame(place_lower)
```


### re sort new unique values
```{r, class.output="scroll-100"}
place_lower_df_order <- place_lower_df$place_lower[order(place_lower_df$place_lower)]
class(place_lower_df_order) # character
place_lower_df_order_df <- data.frame(place_lower_df_order)
class(place_lower_df_order_df) # dataframe
```


### re sort new unique values
```{r, class.output="scroll-100"}
place_lower_df_order <- place_lower_df$place_lower[order(place_lower_df$place_lower)]
```

### create dataframe from new unique values
```{r, class.output="scroll-100"}
class(place_lower_df_order) # character
place_lower_df_order_df <- data.frame(place_lower_df_order)
class(place_lower_df_order_df) # dataframe
```

# Save csv file of unique place names
* unique place names after renaming
```{r, class.output="scroll-100"}
unique_place_after_df_order_df  #56
place_lower_df_order_df  #56

write.csv(unique_place_after_df_order_df, "unique_place_after_df_order_df.csv")
write.csv(place_lower_df_order_df, "place_lower_df_order_df.csv")
```


# Subset dataframe into new place-based objects
```{r, class.output="scroll-100"}
# indexing based on value of place
# uplands <- wa_inat_sel[wa_inat_sel$place == 'Uplands Park', ]
# class(uplands)
### SAME AS
# uplands_park_sub <- subset(wa_inat_sel, place=='Uplands Park')

# subset
albert_head_lagoon_sub <- subset(wa_inat_sel, place=='Albert Head Lagoon')
beacon_hill_park_sub <- subset(wa_inat_sel, place=='Beacon Hill Park')
beaver_lake_ponds_sub <- subset(wa_inat_sel, place=='Beaver Lake Ponds')
botanical_beach_provincial_park_sub <- subset(wa_inat_sel, place=='Botanical Beach Provincial Park')
burnside_sub <- subset(wa_inat_sel, place=='Burnside')
buttle_lake_sub <- subset(wa_inat_sel, place=='Buttle Lake')
cadboro_bay_sub <- subset(wa_inat_sel, place=='Cadboro Bay')
capital_sub <- subset(wa_inat_sel, place=='Capital')
cattle_point_sub <- subset(wa_inat_sel, place=='Cattle Point')
cervus_creek_sub <- subset(wa_inat_sel, place=='Cervus Creek')
china_beach_sub <- subset(wa_inat_sel, place=='China Beach')
coburg_peninsula_sub <- subset(wa_inat_sel, place=='Coburg Peninsula')
coles_bay_regional_park_sub <- subset(wa_inat_sel, place=='Coles Bay Regional Park')
crest_lake_sub <- subset(wa_inat_sel, place=='Crest Lake')
cunningham_woods_sub <- subset(wa_inat_sel, place=='Cunningham Woods')
drum_lake_sub <- subset(wa_inat_sel, place=='Drum Lake')
elk_river_sub <- subset(wa_inat_sel, place=='Elk River')
esquimalt_lagoon_sub <- subset(wa_inat_sel, place=='Esquimalt Lagoon')
fort_rodd_hill_park_sub <- subset(wa_inat_sel, place=='Fort Rodd Hill Park')
french_beach_provincial_park_sub <- subset(wa_inat_sel, place=='French Beach Provincial Park')
goldstream_provincial_park_sub <- subset(wa_inat_sel, place=='Goldstream Provincial Park')
gorge_park_sub <- subset(wa_inat_sel, place=='Gorge Park')
gorge_tillicum_area_sub <- subset(wa_inat_sel, place=='Gorge Tillicum Area')
harling_point_sub <- subset(wa_inat_sel, place=='Harling Point')
island_view_beach_sub <- subset(wa_inat_sel, place=='Island View Beach')
jordan_river_sub <- subset(wa_inat_sel, place=='Jordan River')
juan_de_fuca_trail_sub <- subset(wa_inat_sel, place=='Juan de Fuca Trail')
loss_creek_sub <- subset(wa_inat_sel, place=='Loss Creek')
mary_lake_nature_sanctuary_sub <- subset(wa_inat_sel, place=='Mary Lake Nature Sanctuary')
mckenzie_bight_sub <- subset(wa_inat_sel, place=='McKenzie Bight')
miracle_beach_provincial_park_sub <- subset(wa_inat_sel, place=='Miracle Beach Provincial Park')
mt_tolmie_sub <- subset(wa_inat_sel, place=='Mt Tolmie')
myra_falls_sub <- subset(wa_inat_sel, place=='Myra Falls')
mystic_beach_sub <- subset(wa_inat_sel, place=='Mystic Beach')
oak_bay_sub <- subset(wa_inat_sel, place=='Oak Bay')
ocean_boulevard_sub <- subset(wa_inat_sel, place=='Ocean Boulevard')
otter_point_sub <- subset(wa_inat_sel, place=='Otter Point')
pacific_rim_national_park_sub <- subset(wa_inat_sel, place=='Pacific Rim National Park')
partridge_hills_sub <- subset(wa_inat_sel, place=='Partridge Hills')
patricia_bay_sub <- subset(wa_inat_sel, place=='Patricia Bay')
pete_wolf_creek_sub <- subset(wa_inat_sel, place=='Pete Wolf Creek')
rithets_bog_park_sub <- subset(wa_inat_sel, place=='Rithets Bog Park')
royal_roads_sub <- subset(wa_inat_sel, place=='Royal Roads')
sandcut_beach_sub <- subset(wa_inat_sel, place=='Sandcut Beach')
shirley_sub <- subset(wa_inat_sel, place=='Shirley')
snidcel_sub <- subset(wa_inat_sel, place=='SNIDCEL')
sooke_river_sub <- subset(wa_inat_sel, place=='')
strathcona_provincial_park_sub <- subset(wa_inat_sel, place=='Strathcona Provincial Park')
taylor_beach_sub <- subset(wa_inat_sel, place=='Taylor Beach')
thetis_lake_sub <- subset(wa_inat_sel, place=='Thetis Lake')
university_of_victoria_sub <- subset(wa_inat_sel, place=='University of Victoria')
uplands_park_sub <- subset(wa_inat_sel, place=='Uplands Park')
victoria_sub <- subset(wa_inat_sel, place=='Victoria')
wiffin_spit_sub <- subset(wa_inat_sel, place=='Wiffin Spit')
```


## recheck unique vales
```{r, class.output="scroll-100"}
unique_place_after <- unique(wa_inat_sel$place)
unique_place_after_df <- data.frame(unique_place_after)
```

### re sort new unique values
```{r, class.output="scroll-100"}
unique_place_after_df_order <- unique_place_after_df$unique_place[order(unique_place_after_df$unique_place)]
class(unique_place_after_df_order) # character
```


### create dataframe from new unique place names
```{r, class.output="scroll-100"}
unique_place_after_df_order_df <- data.frame(unique_place_after_df_order)
class(unique_place_after_df_order_df) # dataframe
```


### unique place names after renaming
```{r, class.output="scroll-100"}
unique_place_after_df_order_df  #56
```

# write csv files for each place
* save in csv/ folder
```{r, class.output="scroll-100"}
write.csv(albert_head_lagoon_sub, "csv/albert_head_lagoon_sub.csv")
write.csv(beacon_hill_park_sub, "csv/beacon_hill_park_sub.csv")
write.csv(beaver_lake_ponds_sub, "csv/beaver_lake_ponds_sub.csv")
write.csv(botanical_beach_provincial_park_sub, "csv/botanical_beach_provincial_park_sub.csv")
write.csv(burnside_sub, "csv/burnside_sub.csv")
write.csv(buttle_lake_sub, "csv/buttle_lake_sub.csv")
write.csv(cadboro_bay_sub, "csv/cadboro_bay_sub.csv")
write.csv(capital_sub, "csv/capital_sub.csv")
write.csv(cattle_point_sub, "csv/cattle_point_sub.csv")
write.csv(cervus_creek_sub, "csv/cervus_creek_sub.csv")
write.csv(china_beach_sub, "csv/china_beach_sub.csv")
write.csv(coburg_peninsula_sub, "csv/coburg_peninsula_sub.csv")
write.csv(coles_bay_regional_park_sub, "csv/coles_bay_regional_park_sub.csv")
write.csv(crest_lake_sub, "csv/crest_lake_sub.csv")
write.csv(cunningham_woods_sub, "csv/cunningham_woods_sub.csv")
write.csv(drum_lake_sub, "csv/drum_lake_sub.csv")
write.csv(elk_river_sub, "csv/elk_river_sub.csv")
write.csv(esquimalt_lagoon_sub, "csv/esquimalt_lagoon_sub.csv")
write.csv(fort_rodd_hill_park_sub, "csv/fort_rodd_hill_park_sub.csv")
write.csv(french_beach_provincial_park_sub, "csv/french_beach_provincial_park_sub.csv")
write.csv(goldstream_provincial_park_sub, "csv/goldstream_provincial_park_sub.csv")
write.csv(gorge_park_sub, "csv/gorge_park_sub.csv")
write.csv(gorge_tillicum_area_sub, "csv/gorge_tillicum_area_sub.csv")
write.csv(harling_point_sub, "csv/harling_point_sub.csv")
write.csv(island_view_beach_sub, "csv/island_view_beach_sub.csv")
write.csv(jordan_river_sub, "csv/jordan_river_sub.csv")
write.csv(juan_de_fuca_trail_sub, "csv/juan_de_fuca_trail_sub.csv")
write.csv(loss_creek_sub, "csv/loss_creek_sub.csv")
write.csv(mary_lake_nature_sanctuary_sub, "csv/mary_lake_nature_sanctuary_sub.csv")
write.csv(mckenzie_bight_sub, "csv/mckenzie_bight_sub.csv")
write.csv(miracle_beach_provincial_park_sub, "csv/miracle_beach_provincial_park_sub.csv")
write.csv(mt_tolmie_sub, "csv/mt_tolmie_sub.csv")
write.csv(myra_falls_sub, "csv/myra_falls_sub.csv")
write.csv(mystic_beach_sub, "csv/mystic_beach_sub.csv")
write.csv(oak_bay_sub, "csv/oak_bay_sub.csv")
write.csv(ocean_boulevard_sub, "csv/ocean_boulevard_sub.csv")
write.csv(otter_point_sub, "csv/otter_point_sub.csv")
write.csv(pacific_rim_national_park_sub, "csv/pacific_rim_national_park_sub.csv")
write.csv(partridge_hills_sub, "csv/partridge_hills_sub.csv")
write.csv(patricia_bay_sub, "csv/patricia_bay_sub.csv")
write.csv(pete_wolf_creek_sub, "csv/pete_wolf_creek_sub.csv")
write.csv(rithets_bog_park_sub, "csv/rithets_bog_park_sub.csv")
write.csv(royal_roads_sub, "csv/royal_roads_sub.csv")
write.csv(sandcut_beach_sub, "csv/sandcut_beach_sub.csv")
write.csv(shirley_sub, "csv/shirley_sub.csv")
write.csv(snidcel_sub, "csv/snidcel_sub.csv")
write.csv(sooke_river_sub, "csv/sooke_river_sub.csv")
write.csv(strathcona_provincial_park_sub, "csv/strathcona_provincial_park_sub.csv")
write.csv(taylor_beach_sub, "csv/taylor_beach_sub.csv")
write.csv(thetis_lake_sub, "csv/thetis_lake_sub.csv")
write.csv(university_of_victoria_sub, "csv/university_of_victoria_sub.csv")
write.csv(uplands_park_sub, "csv/uplands_park_sub.csv")
write.csv(victoria_sub, "csv/victoria_sub.csv")
write.csv(wiffin_spit_sub, "csv/wiffin_spit_sub.csv")
```

## Read csv file example
* created objects with same names used in this script
```{r, class.output="scroll-100"}
goldstream_provincial_park_sub <- read.csv("csv/goldstream_provincial_park_sub.csv")
```


### Create ggplot from read csv object
```{r, class.output="scroll-100"}
ggplot(goldstream_provincial_park_sub) +
  aes(x = scientific_name, y = date_time) +
  geom_tile(size = 1) +
#  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
  scale_x_discrete(limits=rev) + # reverses order of labels
  labs(title = "My Goldtream Park iNaturalist Research Grade Observations") +
  #theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 4), axis.text.y = element_text(size = 4)) +
  coord_flip()

getwd()
ggsave("wa_inat_sel_scinam_goldstream.jpg")
```


# Create a loop for unique places
* doesn't work ???
```{r, class.output="scroll-100"}
###############################################
################################################
## DOESN"T WORK
################################################
################################################

### For loop through unique place
## DOESN"T WORK >>> JUST GIVES EMPTY DATAFRAME
unique_places <- unique_place_after_df_order_df

for(i in 1:nrow(unique_places)) {       # for-loop over rows
  place_i <- wa_inat_sel[wa_inat_sel$place == 'i', ]
}
```

---

# FUNCTION for GGPLOT to Replace Place in script
* inat_ggplot_place_species()
```{r, class.output="scroll-100"}
inat_ggplot_place_species <- function(place, titleplace, filename){
    ggplot(place) +
#    aes(x = scientific_name, y = observed_on_string) +  # this one works, but I don't like date format
    # aes(x = scientific_name, y = date_time) +
    aes(x = scientific_name, y = date) +
    geom_tile(size = .1, aes(fill = Genus), colour = "grey50") +
    #  scale_x_discrete(guide = guide_axis(n.dodge=3)) +
    scale_x_discrete(limits = rev) + # reverses order of labels
    labs(subtitle = titleplace, title = "My iNaturalist Research Grade Observations", x = "Scientific Name", y = "Date") + # coord_flip
    #theme_classic() +
    # theme for 90° angle to position date right at axis tick
    # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 4), axis.text.y = element_text(size = 4)) +
    # them hjust centres date at axis tick
    theme(
      axis.text.x = element_text(vjust = 1, hjust = .5, size = 8),
      axis.text.y = element_text(size = 5),
      plot.title = element_text(hjust = 0.5, size = 12),
      plot.subtitle = element_text(hjust = 0.5, size = 12),
      legend.position ="bottom",
      legend.text = element_text(size=6)) +
    coord_flip()

  aspect_ratio <- 2.5
  height <- 7
  ggsave(filename, height = 10 , width = 3 * aspect_ratio)
}
```


## Function usage
```{r, class.output="scroll-100"}
# Useage
# inat_ggplot_place_species(place_dataframe_name, "Place Name", "name_of_file.jpg")
```


## Function replacements to save jpg images of plot
```{r, class.output="scroll-100"}
inat_ggplot_place_species(beacon_hill_park_sub, "Beacon Hill Park", "wa_inat_ggplot_bhp.jpg")
inat_ggplot_place_species(botanical_beach_provincial_park_sub, "Botanical Beach Provincial Park", "wa_inat_ggplot_botbeach.jpg")
inat_ggplot_place_species(beaver_lake_ponds_sub, "Beaver Lake Ponds", "wa_inat_ggplot_blp.jpg")
inat_ggplot_place_species(buttle_lake_sub, "Buttle Lake, Strathcona Park", "wa_inat_ggplot_blsp.jpg")
inat_ggplot_place_species(cattle_point_sub, "Cattle Point, Uplands Park", "wa_inat_ggplot_cpup.jpg")
inat_ggplot_place_species(esquimalt_lagoon_sub, "Esquimalt Lagoon", "wa_inat_ggplot_esqlag.jpg")
inat_ggplot_place_species(goldstream_provincial_park_sub, "Goldstream Provincial Park", "wa_inat_ggplot_gpp.jpg")
inat_ggplot_place_species(pacific_rim_national_park_sub, "Pacific Rim National Park", "wa_inat_ggplot_pacrim.jpg")
inat_ggplot_place_species(strathcona_provincial_park_sub, "Strathcona Provincial Park", "wa_inat_ggplot_spp.jpg")
inat_ggplot_place_species(beacon_hill_park_sub, "Beacon Hill Park", "wa_inat_ggplot_bhp.jpg")
inat_ggplot_place_species(uplands_park_sub, "Uplands Park", "wa_inat_ggplot_up.jpg")
inat_ggplot_place_species(university_of_victoria_sub, "University of Victoria", "wa_inat_ggplot_uvic.jpg")
```


# won't plot the whole db, just shows legend of all Genus
```{r, class.output="scroll-100"}
inat_ggplot_place_species(wa_inat_sel, "55 Natural Places", "wa_inat_ggplot_inat_sel.jpg")
```


---

# Session info
```{r session-info}
# to document specific packages used to run script
sessionInfo()
```
