---
title: "Warming Stripes"
author: "Wendy Anthony"
date: "09/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Based on 2018 code by Dominic Royé for [How to Create Warming Stripes in R](https://www.r-bloggers.com/how-to-create-warming-stripes-in-r/)

# Install Packages
```lubridate```    for dates ymd( )  
```tidyverse```    for ggplot2; stringr str_c( )  
```RColorBrewer``` for colour scheme  

```{r install_packages, include=FALSE, echo=FALSE}
if(!require("lubridate")) install.packages("lubridate")
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("RColorBrewer")) install.packages("RColorBrewer")
```

## Load Libraries

```{r load_libraries, echo=TRUE, include=TRUE}
library("tidyverse")
library("lubridate")
library("RColorBrewer")
```

## Working Directory

```{r working_directory}
dir <- "/Users/wanthony/Documents/R/Climate-Weather"
setwd(dir)
getwd()
```

# DATA
## import .csv file of monthly time-series data 

Historical Data: [Search for monthly Environment Canada Station Data](https://climate.weather.gc.ca/historical_data/search_historic_data_stations_e.html?searchType=stnName&timeframe=1&txtStationName=victoria+Int%27l&searchMethod=contains&optLimit=yearRange&StartYear=1840&EndYear=2021&Year=2021&Month=5&Day=9&selRowPerPage=25)  
* .csv file will need some data cleaning / manipulation before use
* columns have monthly and seasonal values, and  annual temperature value

```{r load_data}
temp_victoria <- read_csv("VictoriaTemp-1941-2012.csv")
```

## get data info

```{r data_info}
class(temp_victoria)
str(temp_victoria)
head(temp_victoria)
```

## Select Columns: Year and Year Average temperature  

```{r select_columns_temp_year}
temp_victoria_yr <- select(temp_victoria,Year,YearAvg)
temp_victoria_yr
```

## rename the temperature column

```{r rename_column}
temp_victoria_yr <- rename(temp_victoria_yr, TempMean=YearAvg)
temp_victoria_yr
```

### Check summary stats to see if data contains any missing values i.e. 999.9 (this would show as the max value)  

```{r summary_stats}
summary(temp_victoria_yr)
```

### Use the ifelse( ) function to replace missing values 999.9 with NA  

```{r missing_values}
temp_victoria_yr <- mutate(temp_victoria_yr, 
                           TempMean = ifelse(TempMean == 999.9, NA, TempMean))
temp_victoria_yr
```

# Convert year variable to a date object for ease of use with ```Lubridate``` and ```ggplot2``` packages  
* tidyverse::str_c() function enables character combinations by specifiying a separator sep="-"  
* lubridate::ymd() function converts date character to Date object

```{r mutate_date}
temp_victoria_yr <- mutate(temp_victoria_yr, 
                           date = str_c(Year, "01-01", sep = "-") %>%
                           ymd())
temp_victoria_yr
```

# Create the Stripes
## Create default theme_minimal() style 

```{r theme_minimal}
theme_strip <- theme_minimal()+
                 theme(axis.text.y = element_blank(),
                       axis.line.y = element_blank(),
                       axis.title = element_blank(),
                       panel.grid.major=element_blank(),
                       legend.title = element_blank(),
                       axis.text.x=element_text(vjust=3),
                       panel.grid.minor=element_blank(),
                       plot.title=element_text(size=14,face="bold")
                       )
```

## Assign colours using RColorBrewer  
[ColorBrewer](https://colorbrewer2.org/)  

```{r color_assignment}
# get names of colours
brewer.pal.info

## assign the colors from RColorBrewer to object col_srip.
col_strip <- brewer.pal(11,"RdBu")

```

# Create graphic with color-bar legend using ggplot2
* data doesn't have a value for Y axis, so dummy value of 1 is used  
* geometry used geom_tile()  
* guide_colorbar(barwidth = 1) controls width of legend color bar  

```{r ggplot_graphic_legend}
## stripes plot with the legend
stplt_lg <- ggplot(temp_victoria_yr,
             aes(x = date, y = 1, fill = TempMean)) +
        geom_tile() +
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0)) +
           scale_y_continuous(expand = c(0, 0)) +
           scale_fill_gradientn(colors = rev(col_strip)) +
             guides(fill = guide_colorbar(barwidth = 1)) +
            labs(title = "Victoria BC 1941-2012",
                caption = "Annual Average Temperature Analysis (Celsius)\nUsing historical climate data from Environment Canada\nhttps://climate.weather.gc.ca/historical_data/search_historic_data_e.html") +
          theme_strip
stplt_lg
```


# Create image files with legend

```{r create_image_files_legend}
# png files have very poor resolution
png("VictoriaClimateStripes-1941-2012-lg.png")
stplt_lg
dev.off()

# pdf files have better resolution
pdf("VictoriaClimateStripes-1941-2012-lg.pdf")
stplt_lg
dev.off()

# tiff files have best resolution, with a larger file-size; specify size and resolution
tiff("VictoriaClimateStripes-1941-2012-lg.tiff", units="in", width=5, height=5, res=300)
stplt_lg
dev.off()
```

# Create graphic stripes only, without a color-bar legend
* geom_tile(show.legend = FALSE)  
* use theme_void() to remove all theme styling  
* NA values can have a different color by adding na.value = "gray70" to scale_fill_gradientn()  

```{r ggplot_graphic_no_legend}
stplt <- ggplot(temp_victoria_yr,
             aes(x = date, y = 1, fill = TempMean)) +
        geom_tile(show.legend = FALSE) +
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0)) +
           scale_y_discrete(expand = c(0, 0)) +
           scale_fill_gradientn(colors = rev(col_strip), na.value = "gray70") +
             theme_void()
stplt
```

# Create image files with no legend

```{r create_image_files_no_legend}
# png files have very poor resolution
png("VictoriaClimateStripes-1941-2012-lg.png")
stplt
dev.off()

# pdf files have better resolution
pdf("VictoriaClimateStripes-1941-2012-lg.pdf")
stplt
dev.off()

# tiff files have best resolution, with a larger file-size; specify size and resolution
tiff("VictoriaClimateStripes-1941-2012-lg.tiff", units="in", width=5, height=5, res=300)
stplt
dev.off()
```
