---
title: "Dashboard Example"
author: "yihui"
date: "January 9, 2020"
output: flexdashboard::flex_dashboard
---

```{r setup, include=FALSE}
# install.packages("flexdashboard")
library("flexdashboard")
```
# Dashboard Layout  
https://bookdown.org/yihui/rmarkdown/layout.html
```{r}
```
"Note that a series of dashes under a line of text is the alternative Markdown syntax for the second-level section header" - equivalent to ## Column 1  
" By default, the second-level sections generate columns on a dashboard, and the third level sections will be stacked vertically inside columns. You do not have to have columns on a dashboard: when all you have are the third-level sections in the source document, they will be stacked vertically as one column in the output."  
"The text of the second-level headers will not be displayed in the output. The second-level headers are for the sole purpose of layout, so the actual content of the headers does not matter at all. By contrast, the first-level and third-level headers will be displayed as titles."  

Column 1
-----------------------------------------------------------------------

### Chart A

```{r}
```


Column 2
-----------------------------------------------------------------------

### Chart B

```{r}
```

### Chart C

```{r}
```

Narrow Column 3 {data-width = 350}
----------------------------------------------------------------
### Chart D

output:
  flexdashboard::flex_dashboard:
    orientation: rows
```{r}
```


