---
title: "Image Colour Palette"
author: "Wendy Anthony"
date: "15/03/2021"
output:
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## `imgpalr`
<https://rdrr.io/cran/imgpalr/>

```{r install_imgpalr}
#install.packages("imgpalr")
library(imgpalr)
#help("imgpalr")
```
### Find working directory

```{r wd}
getwd()
```

# Create herringspawn Sentinel 2 image palette
## Load Image

```{r load_image, results=FALSE}
# use image cropped to desired colour for best results
image_load("imgpalr-hs-1.png")
```

## Create string of palette hex codes

```{r imgpalr}
# does a very coarse pass over image colour, not much differentiation if the desired area of colour is too small >> crop image & try again
imgpal_1 <- image_pal("imgpalr-hs-1.png",
           n = 9,
  type = c("qual", "seq", "div"),
  k = 10,
  bw = c(0, 1),
  brightness = c(0, 1),
  saturation = c(0, 1),
  seq_by = "hsv",
  div_center = "#FFFFFF",
  seed = NULL,
  plot = FALSE,
  labels = TRUE,
  label_size = 1,
  label_color = "#000000",
  keep_asp = TRUE,
  quantize = FALSE
)
```

### Summary info about string of hex colour characters

```{r imgpal_values}
imgpal_1
class(imgpal_1) # "character"
str(imgpal_1) # chr [1:9] 
imgpal_1[3] # value of 3 item of character string
```

## Create dataframe of hex colour names
<https://stat.ethz.ch/pipermail/r-help/2015-April/427663.html>  

```{r hex_colnames_create_dataframe}
clrs <- data.frame(Color=colors(), 
                   hex=rgb(t(col2rgb(colors())),  
                           maxColorValue=255),
                   stringsAsFactors=FALSE)
  str(clrs)
```

## Find name of hex values imported from image using `plotrix` package

```{r x}
# https://stat.ethz.ch/pipermail/r-help/2015-April/427663.html
x <- sapply(imgpal_1, plotrix::color.id)
x
class(x) # list
x[3]
```

## Summary info about string of hex colour names

```{r x_value}
x
class(x) # "character"
str(x) # chr [1:9] 
x[7] # value of 3 item of character string
```

### Create test plot using colours with character string position for colour and name of plot title

```{r colour_test_plot}
# using location of character string from image palette hex results
temp <- c(5,7,6,4,8)
barplot(temp, main="By default")
barplot(temp, col=imgpal_1[1], main=imgpal_1[1]) # 
barplot(temp, col=imgpal_1[2], main=imgpal_1[2]) # 
barplot(temp, col=imgpal_1[3], main=imgpal_1[3]) # 
barplot(temp, col=imgpal_1[4], main=imgpal_1[4]) # 
barplot(temp, col=imgpal_1[5], main=imgpal_1[5]) # 
barplot(temp, col=imgpal_1[6], main=imgpal_1[6]) # 
barplot(temp, col=imgpal_1[7], main=imgpal_1[7]) # 
# barplot(temp, col=imgpal_1[7], main=x[7]) # can't use
barplot(temp, col=imgpal_1[8], main=imgpal_1[8]) # 
barplot(temp, col=imgpal_1[9], main=imgpal_1[9]) # 
barplot(temp, col="#577582", main="With dark blue coloring") # 
```



