---
title: "Xmas Carol Wordclouds"
author: "Wendy Anthony"
date: "`r file.info(knitr::current_input())$ctime`"
output: 
  bookdown::html_document2:
    toc: TRUE
    toc_depth: 4
    toc_float: 
      collapsed: false
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 300px; # for vertical scroll
  max-width: 1500px; # for horizontal scroll
  overflow-y: auto; # for vertical scroll
  overflow-x: auto; # for horizontal scroll  
  background-color: inherit;
}
```


```{r load-package, message=FALSE, echo=FALSE}
# https://spencerschien.info/post/data_viz_how_to/dense_word_clouds/

# install.packages("")
library(tidyverse)
# library(svgparser)
library(grid)
library(wordcloud2) # The {wordcloud2} package provides an HTML5 interface to word clouds
library(wordcloud)
library(tidytext)
library(stopwords)
library(showtext)
library(MetBrewer)  # The {MetBrewer} package provides some great palettes
library(glue)
library(ggtext)
library(cowplot)
```


```{r data, class.output="scroll-100", results='hide', echo=FALSE, message=FALSE}
xmas_titles_df <- read.delim("xmas-lyrics-titles.txt")
colnames(xmas_titles_df)[1] <- "Title"
xmas_titles_df

xmas_titles_french_df <- read.delim("xmas-french-titles.txt", fileEncoding = "utf-8", header = FALSE) # to include first line
colnames(xmas_titles_french_df)[1] <- "Titre"
xmas_titles_french_df
```

# English Christmas Carols Wordclouds 

## 46 English Carol Titles of lyrics used
*(scroll to see whole list)*
```{r titles, class.output="scroll-100", echo=FALSE}
xmas_titles_df
```

```{r load-data-xmas-tunes, class.output="scroll-100", results='hide', echo=FALSE}
xmas_tunes_df <- read.delim("xmas-lyrics.txt")
xmas_tunes_df
str(xmas_tunes_df)
```

```{r load-data-xmas-tunes-french, class.output="scroll-100", results='hide', echo=FALSE}
xmas_tunes_french_df <- read.delim("xmas-french-lyrics.txt")
xmas_tunes_french_df
str(xmas_tunes_french_df)
```

```{r dataprep, class.output="scroll-100", results='hide', echo=FALSE, message=FALSE}
# Get countries for stop words
all_countries <- stopwords_getlanguages(source = "snowball")
# Get stop words for all languages available
stops <- purrr::map_df(all_countries, function(x) get_stopwords(language = x))
class(stops) # [1] "tbl_df"     "tbl"        "data.frame"
str(stops)

# character vectors
# class(fr_sw) # [1] "character"
# same as
# stopwords::data_stopwords_snowball$fr
fr_sw <- stopwords::stopwords("fr", source = "snowball")
lexicon <- rep("lexicon", times = length(fr_sw))
# create df
fr_sw_df <- data.frame(fr_sw, lexicon)
# change column name 
colnames(fr_sw_df)[1] <- "word"
fr_sw_df

# rbind add rows from 2 dataframes
stops <- rbind(stops, fr_sw_df)





# class(stops) # "tbl_df"     "tbl"        "data.frame"

# Get words, remove stop words
xmas_tunes_df_prep <- xmas_tunes_df %>%
  transmute(line = 1:nrow(xmas_tunes_df),
            text = Popular.Christmas.Carol.Lyrics) %>%
  unnest_tokens(word, text) %>%
  anti_join(stops) %>%
  # this final step removes a leading apostrophe in words like l'amour
  mutate(word = str_remove(word, "^\\w'"))
xmas_tunes_df_prep

# Get frequencies by grouping by word and tallying
xmas_tunes_df_tidy <- xmas_tunes_df_prep %>%
  group_by(word) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq))
```


```{r dataprep-french, class.output="scroll-100", results='hide', echo=FALSE, message=FALSE}
# Get countries for stop words
all_countries <- stopwords_getlanguages(source = "snowball")

# Get stop words for all languages available
stops <- map_df(all_countries, function(x) get_stopwords(language = x))

# str(xmas_tunes_french_df)
# Get words, remove stop words
xmas_tunes_french_df_prep <- xmas_tunes_french_df %>%
  transmute(line = 1:nrow(xmas_tunes_french_df),
            text = noel.lyrics) %>%
  unnest_tokens(word, text) %>%
  anti_join(stops) %>%
  # this final step removes a leading apostrophe in words like l'amour
  mutate(word = str_remove(word, "^\\w'"))
xmas_tunes_french_df_prep

# Get frequencies by grouping by word and tallying
xmas_tunes_french_df_tidy <- xmas_tunes_french_df_prep %>%
  group_by(word) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq))
```

```{r explore-data, class.output="scroll-100", results='hide', echo=FALSE}
head(xmas_tunes_df_tidy)
xmas_tunes_df_tidy
```


```{r english_wc, echo=FALSE, results='hide'}
# black_xmas_star_wordcloud_en
# gold_xmas_star_wordcloud_en
# xmas_metbrewer_homer_wordcloud_en
# xmas_default_wordcloud_en
# xmas_red_green_round_wordcloud_en
# xmas_red_green_grey_wordcloud_en
# xmas_red_green_wordcloud_en
# xmas_red_green_purple_wordcloud_en
# xmas_red_green_purple_grey_wordcloud_en
# xmas_red_green_grey_shrt_wordcloud_en
# xmas_red_green_purple_grey_shrt_wordcloud_en
```

## Interactive English Wordclouds
*(Hover over the words to see how often words are used in these 46 carols)*
```{r widget-size, class.output="scroll-100", echo=FALSE}
# wordcloud2(xmas_tunes_df_tidy, shape = "star", 
#            backgroundColor = "black")
# 
# wordcloud2(xmas_tunes_df_tidy, shape = "star", 
#            backgroundColor = "black", gridSize = 3)
# wordcloud2(xmas_tunes_df_tidy, shape = "star", 
#            backgroundColor = "black", widgetsize = c(800,600))
# 
# wordcloud2(xmas_tunes_df_tidy, shape = "star", 
#            backgroundColor = "black", widgetsize = c(1600,800))
# 
# wordcloud2(xmas_tunes_df_tidy, shape = "star", 
#            backgroundColor = "black", widgetsize = c(800,1600))
# 
# wordcloud2(xmas_tunes_df_tidy, shape = "star", 
#            backgroundColor = "black", widgetsize = c(1000,1600))
# 
# wordcloud2(xmas_tunes_df_tidy, shape = "star", 
#            backgroundColor = "black", widgetsize = c(1400,1000))

```
### Black Star
```{r, class.output="scroll-100", echo=FALSE, fig.cap="Black Star"}
black_xmas_star_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, shape = "star", 
           backgroundColor = "black", widgetsize = c(1300,1000))
black_xmas_star_wordcloud_en
```

### Gold Star
```{r, class.output="scroll-100", echo=FALSE, fig.cap="Gold Star"}
gold_xmas_star_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, shape = "star", 
                                       backgroundColor = "gold", size = 0.75,
                                       widgetsize = c(1000,800))
gold_xmas_star_wordcloud_en
# class(gold_xmas_star_wordcloud)
```


```{r, echo=FALSE, fig.cap=""}

```

### MetBrewer: Homer Palette
```{r, echo=FALSE, fig.cap="MetBrewer: Homer Palette"}
colors <- met.brewer("Homer2")
c_trim <- c(rep(colors[c(1,2,3,4,5,6)], 50), rep(colors[length(colors)], nrow(xmas_tunes_df_tidy) - 50))
xmas_metbrewer_homer_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, 
           widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
           color = c_trim, fontFamily = "Julee")
xmas_metbrewer_homer_wordcloud_en
```

### Default shape and colour
```{r, echo=FALSE, fig.cap="Default shape and colour"}
xmas_default_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, widgetsize = c(1000,800))
xmas_default_wordcloud_en
```


### Round red and green
```{r, echo=FALSE, fig.cap="Round red and green"}
colorVec = rep(c('red', 'green'), length.out=nrow(xmas_tunes_df_tidy))
xmas_red_green_round_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, 
           widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
           color = colorVec, fontFamily = "Julee")
xmas_red_green_round_wordcloud_en
```

```{r change-color-size, class.output="scroll-100", echo=FALSE}
## Change colour and size

# The {MetBrewer} package provides some great palettes
# MetBrewer::display_all(sequential = FALSE, colorblind_only = TRUE)
# MetBrewer::display_all(sequential = FALSE, colorblind_only = FALSE)

# Set colors, I wanted pink and blue since I already saw 'love' is tops

# colors <- met.brewer("Homer2")
# c_trim <- c(rep(colors[c(1,2,3,4,5,6)], 50), rep(colors[length(colors)], nrow(xmas_tunes_df_tidy) - 50))
# xmas_metbrewer_homer_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, 
#            widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
#            color = c_trim, fontFamily = "Julee")
# xmas_metbrewer_homer_wordcloud_en

# xmas_default_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, widgetsize = c(1000,800))
# xmas_default_wordcloud_en

# wordcloud2(xmas_tunes_df_tidy, shape = "star", backgroundColor = "black")
# wordcloud2(xmas_tunes_df_tidy, shape = "star", color = colors, backgroundColor = "white")
# wordcloud2(xmas_tunes_df_tidy, shape = "star", color = c_trim, backgroundColor = "green")

# colorVec = rep(c('red', 'green'), length.out=nrow(xmas_tunes_df_tidy))
# xmas_red_green_round_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, 
#            widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
#            color = colorVec, fontFamily = "Julee")
# xmas_red_green_round_wordcloud_en
```


```{r custom-color-vector, class.output="scroll-100", echo=FALSE}
#### Demo using custom color vector

# colorVec = rep(c('red', 'green'), length.out=nrow(demoFreq))
# wordcloud2(xmas_tunes_df_tidy, color = colorVec, fontWeight = "bold")
```


```{r themes, echo=FALSE}
## Themes 

# wc = wordcloud2(xmas_tunes_df_tidy, widgetsize = c(1000,800), )
# wc
# wc + WCtheme(1)
# wc + WCtheme(2)
# wc + WCtheme(3)
# wc + WCtheme(2) + WCtheme(3)
```


```{r themes-star, echo=FALSE}
### Themes - xmas star

# gold_xmas_star_wordcloud
# gold_xmas_star_wordcloud + WCtheme(1)
# gold_xmas_star_wordcloud + WCtheme(2)
# gold_xmas_star_wordcloud + WCtheme(3)
# gold_xmas_star_wordcloud + WCtheme(2) + WCtheme(3)
```


```{r font-weight, echo=FALSE}
## Font weight
# wordcloud2(xmas_tunes_df_tidy, color = colorVec, fontWeight = "bold")
# wordcloud2(xmas_tunes_df_tidy, color = colorVec)
```

### Colour by frequency
#### red, green, grey
```{r ifelse-frequency-color, echo=FALSE, fig.cap="Colour by frequency: red, green, grey"}
# Color by Frequency
# wordcloud2(xmas_tunes_df_tidy,
# color = ifelse(xmas_tunes_df_tidy[, 2] > 20, 'red', 'green'))

xmas_red_green_grey_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, widgetsize = c(1000,800), 
                color = ifelse(xmas_tunes_df_tidy[, 2] < 9, 'grey',
                ifelse(xmas_tunes_df_tidy[, 2] > 20, 'red', 'green')))
xmas_red_green_grey_wordcloud_en
```

```{r new-column-na<10-ifelse-subset, echo=FALSE}
# create new column
xmas_tunes_df_tidy$freq_high <- xmas_tunes_df_tidy$freq
# xmas_tunes_df_tidy

xmas_tunes_df_tidy$freq_high = ifelse(xmas_tunes_df_tidy$freq_high < 6, 0, xmas_tunes_df_tidy$freq_high)
# xmas_tunes_df_tidy

# subset by removing rows with frequency = 0 (i.e. <8)
xmas_tunes_df_tidy_short <- xmas_tunes_df_tidy[xmas_tunes_df_tidy$freq_high != 0, ]
# xmas_tunes_df_tidy_short
```

#### red, green
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green"}
xmas_red_green_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, 
                            widgetsize = c(1000,800), 
                            color = ifelse(xmas_tunes_df_tidy[, 3] > 20, 'red', 'green'))
xmas_red_green_wordcloud_en
```

#### red, green, purple
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green, purple"}
xmas_red_green_purple_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, 
                            widgetsize = c(1000,800), 
                            color = ifelse(xmas_tunes_df_tidy[, 3] > 20, 'red',
                            ifelse(xmas_tunes_df_tidy[, 3] > 0, 'green', 'purple')))
xmas_red_green_purple_wordcloud_en
```

#### red, green, purple, grey
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green, purple, grey"}
xmas_red_green_purple_grey_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy, 
                            widgetsize = c(1000,800), 
                            color = ifelse(xmas_tunes_df_tidy[, 2] >  20, 'red',
                            ifelse(xmas_tunes_df_tidy[, 2] > 8, 'purple',
                            ifelse(xmas_tunes_df_tidy[, 2] > 3, 'green', 'grey'))))
xmas_red_green_purple_grey_wordcloud_en
```

#### red, green, grey - frequency >2
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green, grey - frequency >2"}
xmas_red_green_grey_shrt_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy_short, 
                          widgetsize = c(1000,800), 
                          color = ifelse(xmas_tunes_df_tidy_short[, 3] < 13, 'grey',
                          ifelse(xmas_tunes_df_tidy_short[, 3] > 20, 'red', 'green')))
xmas_red_green_grey_shrt_wordcloud_en
```

#### red, green - frequency >2
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green - frequency >2"}
xmas_red_green_purple_grey_shrt_wordcloud_en <- wordcloud2(xmas_tunes_df_tidy_short,
                          widgetsize = c(1000,800), 
                          color = ifelse(xmas_tunes_df_tidy_short[, 3] >  20, 'red',
                          ifelse(xmas_tunes_df_tidy_short[, 3] > 8, 'green',
                          ifelse(xmas_tunes_df_tidy_short[, 3] > 7, 'purple', 'grey'))))
xmas_red_green_purple_grey_shrt_wordcloud_en
```


---

# French Christmas Carols Wordclouds

## 15 FrenchCarol Titles of lyrics used
*(scroll to see whole list)*
```{r french-titles, class.output="scroll-100", echo=FALSE}
xmas_titles_french_df
```

## Interactive French Wordclouds
*(Hover over the words to see how often words are used in these 15 carols)*

```{r frequency-column, echo=FALSE}
xmas_tunes_french_df_tidy$freq_high <- xmas_tunes_french_df_tidy$freq
# xmas_tunes_french_df_tidy

xmas_tunes_french_df_tidy$freq_high = ifelse(xmas_tunes_french_df_tidy$freq_high < 2, 0, xmas_tunes_french_df_tidy$freq_high)
# xmas_tunes_french_df_tidy

# subset by removing rows with frequency = 0 (i.e. <8)
xmas_tunes_french_df_tidy_short <- xmas_tunes_french_df_tidy[xmas_tunes_french_df_tidy$freq_high != 0, ]
# xmas_tunes_french_df_tidy_short

```

### Black Star
```{r, echo=FALSE, fig.cap="Black Star"}
black_xmas_star_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, shape = "star", 
           backgroundColor = "black", widgetsize = c(1300,1000))
black_xmas_star_wordcloud_fr
```

### Gold Star 
```{r, echo=FALSE, fig.cap="Gold Star"}
gold_xmas_star_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, shape = "star", 
                                       backgroundColor = "gold", size = 0.75,
                                       widgetsize = c(1000,800))
gold_xmas_star_wordcloud_fr
```

### MetBrewer: Homer Palette
```{r, echo=FALSE, fig.cap="MetBrewer: Homer Palette"}
colors <- met.brewer("Homer2")
c_trim <- c(rep(colors[c(1,2,3,4,5,6)], 50), rep(colors[length(colors)], nrow(xmas_tunes_french_df_tidy) - 50))
xmas_metbrewer_homer_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
           widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
           color = c_trim, fontFamily = "Julee")
xmas_metbrewer_homer_wordcloud_fr
```

### Default shape and colour 
```{r, echo=FALSE, fig.cap="Default shape and colour"}
xmas_default_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, widgetsize = c(1000,800))
xmas_default_wordcloud_fr
```

### Round red and green
```{r, echo=FALSE, fig.cap=""}
colorVec = rep(c('red', 'green'), length.out=nrow(xmas_tunes_french_df_tidy))
xmas_red_green_round_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
           widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
           color = colorVec, fontFamily = "Julee")
xmas_red_green_round_wordcloud_fr
```

### Colour by frequency
#### red, green, grey
```{r, echo=FALSE, fig.cap="Colour by frequency: "}
xmas_red_green_grey_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, widgetsize = c(1000,800),                          color = ifelse(xmas_tunes_french_df_tidy[, 2] < 9, 'grey',
                         ifelse(xmas_tunes_french_df_tidy[, 2] > 20, 'red', 'green')))
xmas_red_green_grey_wordcloud_fr
```

#### red, green
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green"}
xmas_red_green_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, widgetsize = c(1000,800), 
                            color = ifelse(xmas_tunes_french_df_tidy[, 3] > 20, 'red', 'green'))
xmas_red_green_wordcloud_fr
```

#### red, green, purple
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green, purple"}
xmas_red_green_purple_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
                        widgetsize = c(1000,800), 
                        color = ifelse(xmas_tunes_french_df_tidy[, 3] > 20, 'red',
                        ifelse(xmas_tunes_french_df_tidy[, 3] > 0, 'green', 'purple')))
xmas_red_green_purple_wordcloud_fr
```

#### red, green, purple, grey
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green, purple, grey"}
xmas_red_green_purple_grey_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
                        widgetsize = c(1000,800), 
                        color = ifelse(xmas_tunes_french_df_tidy[, 2] >  20, 'red',
                        ifelse(xmas_tunes_french_df_tidy[, 2] > 8, 'purple',
                        ifelse(xmas_tunes_french_df_tidy[, 2] > 3, 'green', 'grey'))))
xmas_red_green_purple_grey_wordcloud_fr
```

#### red, green, grey - frequency >2
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green, grey - frequency >2"}
xmas_red_green_grey_shrt_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy_short, 
                        widgetsize = c(1000,800), 
                        color = ifelse(xmas_tunes_french_df_tidy_short[, 3] < 13, 'grey',
                        ifelse(xmas_tunes_french_df_tidy_short[, 3] > 20, 'red', 'green')))
xmas_red_green_grey_shrt_wordcloud_fr
```

#### red, green - frequency >2
```{r, echo=FALSE, fig.cap="Colour by frequency: red, green - frequency >2"}
xmas_red_green_purple_grey_shrt_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy_short,
                        widgetsize = c(1000,800), 
                        color = ifelse(xmas_tunes_french_df_tidy_short[, 3] >  20, 'red',
                        ifelse(xmas_tunes_french_df_tidy_short[, 3] > 5, 'green',
                        ifelse(xmas_tunes_french_df_tidy_short[, 3] > 2, 'purple', 'grey'))))
xmas_red_green_purple_grey_shrt_wordcloud_fr
```

```{r french-interactive, echo=FALSE}
# black_xmas_star_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, shape = "star", 
#            backgroundColor = "black", widgetsize = c(1300,1000))
# black_xmas_star_wordcloud_fr

# gold_xmas_star_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, shape = "star", 
#                                        backgroundColor = "gold", size = 0.75,
#                                        widgetsize = c(1000,800))
# gold_xmas_star_wordcloud_fr

# colors <- met.brewer("Homer2")
# c_trim <- c(rep(colors[c(1,2,3,4,5,6)], 50), rep(colors[length(colors)], nrow(xmas_tunes_french_df_tidy) - 50))
# xmas_metbrewer_homer_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
#            widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
#            color = c_trim, fontFamily = "Julee")
# xmas_metbrewer_homer_wordcloud_fr

# xmas_default_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, widgetsize = c(1000,800))
# xmas_default_wordcloud_fr

# colorVec = rep(c('red', 'green'), length.out=nrow(xmas_tunes_french_df_tidy))
# xmas_red_green_round_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
#            widgetsize = c(1000,800), ellipticity = .9, gridSize = 10,
#            color = colorVec, fontFamily = "Julee")
# xmas_red_green_round_wordcloud_fr

# xmas_red_green_grey_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, widgetsize = c(1000,800),                          color = ifelse(xmas_tunes_french_df_tidy[, 2] < 9, 'grey',
#                          ifelse(xmas_tunes_french_df_tidy[, 2] > 20, 'red', 'green')))
# xmas_red_green_grey_wordcloud_fr

# xmas_tunes_french_df_tidy$freq_high <- xmas_tunes_french_df_tidy$freq
# # xmas_tunes_french_df_tidy
# 
# xmas_tunes_french_df_tidy$freq_high = ifelse(xmas_tunes_french_df_tidy$freq_high < 2, 0, xmas_tunes_french_df_tidy$freq_high)
# # xmas_tunes_french_df_tidy

# xmas_red_green_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, widgetsize = c(1000,800), 
#                             color = ifelse(xmas_tunes_french_df_tidy[, 3] > 20, 'red', 'green'))
# xmas_red_green_wordcloud_fr

# xmas_red_green_purple_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
#                         widgetsize = c(1000,800), 
#                         color = ifelse(xmas_tunes_french_df_tidy[, 3] > 20, 'red',
#                         ifelse(xmas_tunes_french_df_tidy[, 3] > 0, 'green', 'purple')))
# xmas_red_green_purple_wordcloud_fr

# xmas_red_green_purple_grey_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy, 
#                         widgetsize = c(1000,800), 
#                         color = ifelse(xmas_tunes_french_df_tidy[, 2] >  20, 'red',
#                         ifelse(xmas_tunes_french_df_tidy[, 2] > 8, 'purple',
#                         ifelse(xmas_tunes_french_df_tidy[, 2] > 3, 'green', 'grey'))))
# xmas_red_green_purple_grey_wordcloud_fr

# # subset by removing rows with frequency = 0 (i.e. <8)
# xmas_tunes_french_df_tidy_short <- xmas_tunes_french_df_tidy[xmas_tunes_french_df_tidy$freq_high != 0, ]
# # xmas_tunes_french_df_tidy_short

# xmas_red_green_grey_shrt_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy_short, 
#                         widgetsize = c(1000,800), 
#                         color = ifelse(xmas_tunes_french_df_tidy_short[, 3] < 13, 'grey',
#                         ifelse(xmas_tunes_french_df_tidy_short[, 3] > 20, 'red', 'green')))
# xmas_red_green_grey_shrt_wordcloud_fr

# xmas_red_green_purple_grey_shrt_wordcloud_fr <- wordcloud2(xmas_tunes_french_df_tidy_short,
#                         widgetsize = c(1000,800), 
#                         color = ifelse(xmas_tunes_french_df_tidy_short[, 3] >  20, 'red',
#                         ifelse(xmas_tunes_french_df_tidy_short[, 3] > 5, 'green',
#                         ifelse(xmas_tunes_french_df_tidy_short[, 3] > 2, 'purple', 'grey'))))
# xmas_red_green_purple_grey_shrt_wordcloud_fr
```

```{r french_wc, echo=FALSE, results='hide'}
# black_xmas_star_wordcloud_fr
# gold_xmas_star_wordcloud_fr
# xmas_metbrewer_homer_wordcloud_fr
# xmas_default_wordcloud_fr
# xmas_red_green_round_wordcloud_fr
# xmas_red_green_grey_wordcloud_fr
# xmas_red_green_wordcloud_fr
# xmas_red_green_purple_wordcloud_fr
# xmas_red_green_purple_grey_wordcloud_fr
# xmas_red_green_grey_shrt_wordcloud_fr
# xmas_red_green_purple_grey_shrt_wordcloud_fr
```


---


```{r, echo=FALSE}
# Annimation
# install.packages("gifski")
# library(gifski)
```


---


```{r session-info, echo=FALSE, results='hide'}
# to document specific packages used to run script
sessionInfo()
```
