---
title: "Showcase Shiny GOE Monitor DataViz"
author: "Wendy Anthony"
date: "2025/08/2y"
output:
  html_document:
    toc: yes 
    toc_depth: 3
    toc_float: yes
    number_sections: yes
---
Creation date: 2025/08/27
Modified Date & Time: 2025/08/27 @ 08:02

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r load_libraries}
library(shiny)
library(tidyverse)
library(plotly)
```
</p></details>


# Data
- Note Data is in `data_raw/TEST-GOE-Monitor.xlsx` file, with one worksheet for Data, and one worksheet for Site info
- Save each worksheet to 2 separate csv files `data/Site-Data-TEST-GOE-Monitor.csv` & `data/SiteData-TEST-GOE-Monitor.csv`

## Load Data
- Data Cleaning to ensure no space at beginning of any text values and column headings
Note: data/SiteDetails-TEST-GOE-Monitor.csv has space in front of values 
 NEED TO change `data_raw/TEST-GOE-Monitor.xlsx` excel file first, then resave as csv
 <details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r load_data}
# Site Data
geo_eccc <- read.csv("data/SiteData-TEST-GOE-Monitor.csv", header = TRUE, sep = ",")
# Site Details
Sites_eccc <- read.csv("data/SiteDetails-TEST-GOE-Monitor.csv", header = TRUE, sep = ",")
```
</p></details>

## Data Discovery

## Column Names Data
 <details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r data_1}
colnames(geo_eccc)
```
</p></details>

## Column Names Sites
 <details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r data_2}
colnames(Sites_eccc)
```
</p></details>

# Data Manipulation

## proportion to Percentage_of_native_species
 <details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r create_percentage_number}
# using scales creates % character instead of number
# Percentage_of_native_species <- scales::percent(geo_eccc$Proportion_of_native_species)
# Percentage_of_native_species 

Percentage_ns <- geo_eccc$Proportion_of_native_species * 100
Percentage_ns
```
</p></details>


## add `Percentage_ns` data to column
<details><summary><b><i>Click to see results of `data`</i></b></summary><p>
```{r add_column}
# https://sparkbyexamples.com/r-programming/add-column-to-dataframe-in-r/
# geo_eccc_1 <- cbind(geo_eccc, Percentage_of_native_species)
# geo_eccc_1

geo_eccc <- cbind(geo_eccc, Percentage_ns)
geo_eccc
```

</p></details>

## Check data structure of new column
 <details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r data_structure_new_column}
str(geo_eccc)
```
</p></details>

## Data Structure Data
- make sure each column is same data type in both data files
 <details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r data_structure_1}
str(geo_eccc)
```
</p></details>

## Data Structure Sites 
- make sure each column is same data type
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r data_structure_2}
str(Sites_eccc)
```
</p></details>

## Combine Dataframes 
- first make sure they have the same number of rows 
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r cbind}
length(unique(geo_eccc$Site))
length(unique(Sites_eccc$Site))
```
</p></details>

### Arrange Order
- Site details Site names are in different order than Site Data, so put each site in alphabetical order first
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r arange_order_df}
# tidyverse
# https://stackoverflow.com/questions/68989228/sort-dataframe-by-column-value-r
Sites_eccc_1 <- Sites_eccc %>% arrange(Sites_eccc$Site)
geo_eccc_1 <- geo_eccc %>% arrange(geo_eccc$Site)
```
</p></details>

### cbind arranged df works
- (`merge` doesn't work, leaves columns empty)

<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r cbind_arranged_df}
geo_eccc_sites <- cbind(Sites_eccc_1, geo_eccc_1)
names(geo_eccc_sites)
```
</p></details>

### remove duplicate columns
- now there are 2 columns with same name for, Subregion, Site
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r remove_duplicate_columns}
# https://stackoverflow.com/questions/7072159/how-do-you-remove-columns-from-a-data-frame
geo_eccc_sites_1 <- geo_eccc_sites[c(-9,-10)]
names(geo_eccc_sites_1)
```
</p></details>

### Rename columns with . using base R
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r rename_columns_baseR}

names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Land.Manager"] <- "Land_Manager"
names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Main.Restoration.Type"] <- "Main_Restoration_Type"
names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Restoration.Intensity"] <- "Restoration_Intensity"
names(geo_eccc_sites_1)[names(geo_eccc_sites_1) == "Area.ha"] <- "Area_ha"

names(geo_eccc_sites_1)
```
</p></details>


### Create and save new csv file with both datasets 
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r create_csv_combined_datasets}
write.csv(geo_eccc_sites_1, "data/geo_eccc_all_site_data.csv", row.names = FALSE)
```
</p></details>


---

# Plots
## Site by Percentage Native Species

```{r plots_site_native_species}
site_natsp <- ggplot(geo_eccc, 
                    aes(x = Site, y = Percentage_ns)) +
                    geom_point(shape = 18, size = 2) +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Site by Percentage Native Species',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-04")
site_natsp

```

## Site by Composite_Index
```{r plots_Composite_Index}
site_natsp <- ggplot(geo_eccc, 
                    aes(x = Site, y = Composite_Index)) +
                    geom_point(shape = 18, size = 2) +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Site by Composite Index',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-04")
site_natsp
```

## ORDER WORKs Point Plot Site by Restoration_Intensity
Need to reorder Restoration_Intensity using forcats library from tidyverse
https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html
```{r plots_site_Restoration_Intensity_order_point}
names(geo_eccc_sites_1)
# Reorder following a precise order
# https://stackoverflow.com/questions/18413756/re-ordering-factor-levels-in-data-frame
str(geo_eccc_sites_1)

# Step 1
desired_levels <- c("high", "moderate", "low", "minimal")
# create new df to create desired levels
geo_eccc_sites_2 <- geo_eccc_sites_1
# this changes data type from character to factor
geo_eccc_sites_2$Restoration_Intensity <- factor(geo_eccc_sites_1$Restoration_Intensity, levels = desired_levels)
print(levels(geo_eccc_sites_2$Restoration_Intensity)) # New order of levels
str(geo_eccc_sites_2)
# step 2
geo_eccc_sites_2 <- geo_eccc_sites_2[order(geo_eccc_sites_2$Restoration_Intensity),]

# values in order now -- ?? how to change to reverse order 
site_restInt <- ggplot(geo_eccc_sites_2,
       aes(x = Site, y = reorder(Restoration_Intensity, desc(Restoration_Intensity)), color = Subregion)) +
  
                    geom_point(shape = 18, size = 2) +
  #geom_bar(stat="identity") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="bottom") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Sites by Restoration Intensity',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-22", x = "Site", y = "Restoration Intensity")

site_restInt


ggplotly(site_restInt)
```

## Change x axis ORDER WORKs Point Plot Site by Restoration_Intensity

```{r plots_site_Restoration_Intensity_order_point_1}
names(geo_eccc_sites_1)
# Reorder following a precise order
str(geo_eccc_sites_1)

# Step 1
desired_levels <- c("high", "moderate", "low", "minimal")
# create new df to create desired levels
geo_eccc_sites_2 <- geo_eccc_sites_1
# this changes data type from character to factor
geo_eccc_sites_2$Restoration_Intensity <- factor(geo_eccc_sites_1$Restoration_Intensity, levels = desired_levels)
print(levels(geo_eccc_sites_2$Restoration_Intensity)) # New order of levels
str(geo_eccc_sites_2)
# step 2
geo_eccc_sites_2 <- geo_eccc_sites_2[order(geo_eccc_sites_2$Restoration_Intensity),]

# values in order now -- ?? how to change to reverse order 
site_restInt_x <- ggplot(geo_eccc_sites_2,
       aes(x = reorder(Site, 
                       desc(Restoration_Intensity)), 
           y = reorder(Restoration_Intensity, desc(Restoration_Intensity)), color = Subregion)) +
  
                    geom_point(shape = 18, size = 2) +
  #geom_bar(stat="identity") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="bottom") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Site by Restoration_Intensity',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-26",
                      x = "Site", y = "Restoration_Intensity")
site_restInt_x


ggplotly(site_restInt_x)
```

## BAR Plot WORKS Site by Restoration_Intensity
```{r plots_site_Restoration_Intensity_order_bar}
names(geo_eccc_sites_1)
# Reorder following a precise order
str(geo_eccc_sites_1)

# Step 1
desired_levels <- c("high", "moderate", "low", "minimal")
# this changes data type from character to factor
geo_eccc_sites_2$Restoration_Intensity <- factor(geo_eccc_sites_1$Restoration_Intensity, levels = desired_levels)
print(levels(geo_eccc_sites_2$Restoration_Intensity)) # New order of levels
str(geo_eccc_sites_2)
# step 2
geo_eccc_sites_2 <- geo_eccc_sites_2[order(geo_eccc_sites_2$Restoration_Intensity),]

# values in order now -- ?? how to change to reverse order 
site_restInt_bar <- ggplot(geo_eccc_sites_2,
       aes(y = reorder(Restoration_Intensity, desc(Restoration_Intensity)), color = Site, fill = Site)) +
                    geom_bar() +
  #geom_bar(stat="identity") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="bottom") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Site by Restoration_Intensity',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-26", y = "Restoration_Intensity")
site_restInt_bar


ggplotly(site_restInt_bar)
```

## Violin Subregion by Exotic_species
```{r plots_Exotic_species}
subregion_exsp <- ggplot(geo_eccc, 
                    aes(x = Subregion, y = Exotic_species)) +
                    geom_violin(fill = "seagreen2") +
                    geom_boxplot(width = 0.1, fill = "sandybrown") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(vjust = 1, size = 9)) +
                    labs(title='Plotting Subregion by Exotic_species',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-04")
subregion_exsp
# plotly won't work here
```

## Violin Subregion by Percentage_ns
```{r plots_Percentage_ns}
names(geo_eccc)
subregion_nsp <- ggplot(geo_eccc, 
                    aes(x = Subregion, y = Percentage_ns)) +
                    geom_violin(fill = "seagreen2") +
                    geom_boxplot(width = 0.1, fill = "sandybrown") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(vjust = 1, size = 9)) +
                    labs(title='Plotting Subregion by Percentage of Native Species',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-27")
subregion_nsp
# plotly won't work here
```

## Cowplot Violin Subregion plot_grid
```{r cowplot_violin_subr}
library(cowplot)

subregion_exsp_1 <- subregion_exsp + labs(title='Exotic Species',
                      subtitle='ECCC',
                      caption = "2025-08-27")

subregion_nsp_1 <- subregion_nsp + labs(title='Native Species',
                      subtitle='ECCC',
                      caption = "2025-08-27")

plot_grid(subregion_exsp_1, subregion_nsp_1, labels = "AUTO")

```


## Subregion by Proportion_of_native_species
```{r plots_Subregion_Percentage_of_native_species}
subregion_natsp <- ggplot(geo_eccc_1, 
                    aes(x = Subregion, y = Percentage_ns)) +
                    geom_violin(fill = "seagreen2") +
                    geom_boxplot(width = 0.1, fill = "sandybrown") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Subregion by Proportion_of_native_species',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-04")
subregion_natsp
```



---
---


# Plots by Separate Sites
- use pivot_longer() for single row with multiple columns convert to long

## pivot longer one row to long df
<details><summary><b><i>Click to see `details`</i></b></summary><p>
```{r pivot_longer}
colnames(geo_eccc_1)
# choose index row from data sorted alphabetically
# delete first 2 columns as they can't be characters ? also Proportion to leave Percentage 
# delete year column

# Anniversary_Island
Anniversary_Island <- geo_eccc_1[1, ]
Anniversary_Island
Anniversary_Island <- Anniversary_Island[, -c(1,2,9)]
Anniversary_Island_long <- Anniversary_Island %>% pivot_longer(everything(), names_to = "variable", values_to = "value")
Anniversary_Island_long

# Gonzales
Gonzales <- geo_eccc_1[15, ]
Gonzales
Gonzales <- Gonzales[, -c(1,2,9)]
Gonzales
Gonzales_long <- Gonzales %>% pivot_longer(everything(), names_to = "variable", values_to = "value")
Gonzales_long

# Uplands
Uplands <- geo_eccc_1[24, ]
Uplands
Uplands <- Uplands[, -c(1,2,9)]
Uplands
Uplands_long <- Uplands %>% pivot_longer(everything(), names_to = "variable", values_to = "value")
Uplands_long
```
</p></details>

## Point Plot Long Uplands
```{r plots_Uplands}
Uplands_long_plot  <-  ggplot(Uplands_long, 
                    aes(x = variable, y = value)) +
  # https://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations
                    ylim(0, 100) +
                    geom_point() +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Uplands Variables',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-04")
Uplands_long_plot
```

## Point Plot to Compare Gonzales Variables
```{r plots_Gonzales_compare_1}
Gonzales_long_plot  <-  ggplot(Gonzales_long, 
                    aes(x = variable, y = value)) +
  # https://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations
                    ylim(0, 100) +
                    geom_point() +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Gonzales Variables',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-04")
Gonzales_long_plot
```

## Point Plot Long Anniversary_Island_long
```{r plots_Gonzales}
Anniversary_Island_long_plot  <-  ggplot(Anniversary_Island_long, 
                    aes(x = variable, y = value)) +
  # https://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations
                    ylim(0, 100) +
                    geom_point() +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Plotting Anniversary Island Variables',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-04")
Anniversary_Island_long_plot
```


## Bar Variables Compare Uplands
```{r compare_bar_Variables_uplands}

compare_bar_uplands  <-  ggplot(Uplands_long, 
                    aes(x = variable, y = value)) +
  
# Exotic_species
  # https://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations
                    ylim(0, 100) +
                    geom_bar(stat = "identity", fill = "seagreen") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Comparinging Uplands Variables',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-08")
compare_bar_uplands
```

## Compare Bar Variables multiple sites
```{r compare_bar_Variables_multiple_sites}

names(geo_eccc_1)
compare_bar_multiple_sites  <-  ggplot(geo_eccc_1,
                    aes(x = Site, y = Percentage_ns, fill = Subregion, color = Subregion)) +
  
# Exotic_species
  # https://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations
                    ylim(0, 100) +
                    geom_bar(stat = "identity") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Comparinging Uplands Variables',
                      subtitle='ECCC',
                      caption = "Chart by Wendy Anthony \n 2025-08-08")
compare_bar_multiple_sites
```

## Arrange Compare Bar Variables multiple sites
```{r compare_bar_Variables_multiple_sites_order}
# https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html
names(geo_eccc_1)
# arrange values
#str(arr)
# geo_eccc_1 %>% arrange(Percentage_ns) -> arr
geo_eccc_1 %>% arrange(desc(Percentage_ns)) %>%
  # updates Site with new arrangement
   mutate(Site = factor(Site, levels = Site)) %>% 
    ggplot(
      aes(x = Site, y = Percentage_ns, fill = Subregion, color = Subregion)) +
  # https://www.sthda.com/english/wiki/ggplot2-axis-scales-and-transformations
                    ylim(0, 100) +
                    geom_bar(stat = "identity") +
                    theme_minimal() + #get rid of grey background and tick marks 
                    theme(legend.position="bottom") + 
                    theme(axis.text.x = element_text(angle = 35, vjust = 1, hjust=1, size = 7)) +
                    labs(title='Comparinging Sites',
                      subtitle='ECCC GOE Site Monitoring',
                      caption = "Chart by Wendy Anthony \n 2025-08-19")

```


---

# Test GGplot Function
## Violin Chart Function
```{r Test_GGplot_Function_violin}
geo_eccc <- read.csv("data/SiteData-TEST-GOE-Monitor.csv", header = TRUE, sep = ",", stringsAsFactors=FALSE)
str(geo_eccc)
names(geo_eccc)

subregion_var_violin <- function(file, title, subtitle, colx, coly, xlab, ylab, caption){
  
    ggplot2::ggplot(file,
        ggplot2::aes(x = colx, y = coly)) +
    ggplot2::geom_violin(fill = "seagreen2") +
    ggplot2::geom_boxplot(width = 0.1, fill = "sandybrown") +
    ggplot2::theme_minimal() + #get rid of grey background and tick marks
    ggplot2::theme(legend.position="none") + #remove legend
    ggplot2::theme(
      axis.text.x = ggplot2::element_text(vjust = 1, hjust = .5, size = 8),
      axis.text.y = ggplot2::element_text(size = 5),
      plot.title = ggplot2::element_text(hjust = 0.5, size = 12),
      plot.subtitle = ggplot2::element_text(hjust = 0.5, size = 12)) +
    #theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
    ggplot2::labs(title=title,
         subtitle=subtitle,
         caption = caption,
         x = xlab, y = ylab)
  #subregion_exsp
  
}

subregion_var_violin(geo_eccc, 'Plotting Subregion by Exotic_species', 'ECCC Data Paper', geo_eccc$Subregion, geo_eccc$Exotic_species, "Subregion", "Exotic_species", "Chart by Wendy Anthony \n 2025-08-04")
subregion_var_violin(geo_eccc,'Plotting Subregion by Cultural_species_richness', 'ECCC Data Paper', geo_eccc$Subregion, geo_eccc$Cultural_species_richness, "Subregion", "Cultural_species_richness", "Chart by Wendy Anthony \n 2025-08-16")

```

## Point Chart Function
### ggplot point chart
```{r Test_GGplot_Function_point}
geo_eccc <- read.csv("data/SiteData-TEST-GOE-Monitor.csv", header = TRUE, sep = ",", stringsAsFactors=FALSE)

str(geo_eccc)
names(geo_eccc)


subregion_var_point <- function(file, title, subtitle, colx, coly, xlab, ylab, caption){

ggplot(file, 
                    aes(x = colx, y = coly, colour = Subregion)) +
                    geom_point(shape = 18, size = 2) +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle =35, vjust = 1, hjust=1, size = 7)) +
                    labs(title=title,
                      subtitle=subtitle,
                      caption = caption,
                      x = xlab, y = ylab)
}

subregion_var_point(geo_eccc,'Plotting Site by Percentage Native Species', 'Malloff & Shackelford. (2024). \nFeeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. \nRestoration Futures Lab at the University of Victoria.
', geo_eccc$Site, geo_eccc$Proportion_of_native_species, "Site", "Proportion of native species", "Chart by Wendy Anthony \n 2025-09-16 \n ECCC Data: (Malloff & Shackelford, 2024)")

subregion_var_point(geo_eccc,'Plotting Site by Exotic Species', 'Malloff & Shackelford. (2024). \nFeeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. \nRestoration Futures Lab at the University of Victoria.
', geo_eccc$Site, geo_eccc$Exotic_species, "Site", "Exotic Species", "Chart by Wendy Anthony \n 2025-09-16 \n ECCC Data: (Malloff & Shackelford, 2024)")
```

### plotly ggplot point chart
```{r Test_GGplot_Function_point_plotly}
geo_eccc <- read.csv("data/SiteData-TEST-GOE-Monitor.csv", header = TRUE, sep = ",", stringsAsFactors=FALSE)

str(geo_eccc)
names(geo_eccc)


subregion_var_point <- function(file, title, subtitle, colx, coly, xlab, ylab, caption){

ggplot(file, 
                    aes(x = colx, y = coly, colour = Subregion)) +
                    geom_point(shape = 18, size = 2) +
                    theme_minimal() + #get rid of grey background and tick marks
                    theme(legend.position="bottom") +
                    # theme(legend.position="none") + #remove legend 
                    theme(axis.text.x = element_text(angle =35, vjust = 1, hjust=1, size = 7)) +
                    labs(title=title,
                      subtitle=subtitle,
                      caption = caption,
                      x = xlab, y = ylab)
}

p2 <- subregion_var_point(geo_eccc,'Plotting Site by Percentage Native Species', 'Malloff & Shackelford. (2024). \nFeeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. \nRestoration Futures Lab at the University of Victoria.
', geo_eccc$Site, geo_eccc$Proportion_of_native_species, "Site", "Proportion of native species", "Chart by Wendy Anthony \n 2025-09-16 \n ECCC Data: (Malloff & Shackelford, 2024)")

subregion_var_point(geo_eccc,'Plotting Site by Exotic Species', 'Malloff & Shackelford. (2024). \nFeeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. \nRestoration Futures Lab at the University of Victoria.
', geo_eccc$Site, geo_eccc$Exotic_species, "Site", "Exotic Species", "Chart by Wendy Anthony \n 2025-09-16 \n ECCC Data: (Malloff & Shackelford, 2024)")

ggplotly(p2)

```


# Leaflet Map
## Custom Icons
```{r leaflet_map_custom_icons}

Sites_eccc <- read.csv("data/SiteDetails-TEST-GOE-Monitor.csv", header = TRUE, sep = ",")

names(Sites_eccc)

library(dplyr)
library(leaflet)
library(htmlwidgets) # for responsive web map

# custom icon
# https://leafletjs.com/examples/custom-icons/
greenLeafIcon <- makeIcon(
  iconUrl = "https://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "https://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)
# change LeafIcon size 1/2
greenLeafIconSm <- makeIcon(
  iconUrl = "https://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 19, iconHeight = 47,
  iconAnchorX = 11, iconAnchorY = 47,
  shadowUrl = "https://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 25, shadowHeight = 32,
  shadowAnchorX = 2, shadowAnchorY = 31
)

# Title control
title <- tags$div(HTML('<b>ECC Site Maps</b><br>Malloff & Shackelford. (2024). Feeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. Restoration Futures Lab at the University of Victoria'))
  

ECCC_map <- leaflet(Sites_eccc) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addMarkers(
             ~ Lng, ~ Lat, icon = greenLeafIconSm,
             popup = paste0("<b>Subregion:</b> ", Sites_eccc$Subregion, "<br>", "<b>Site:</b> ", Sites_eccc$Site, "<br>", "<b>Latitude:</b> ", Sites_eccc$Lat, "<br>", "<b>Longitude:</b> ", Sites_eccc$Lng,  "<br><br>", "<b>Land Manager:</b> ", Sites_eccc$Land.Manager, "<br>", "<b>Main Restoration:</b> ", Sites_eccc$Main.Restoration.Type, "<br>", "<b>Restoration Intensity:</b> ", Sites_eccc$Restoration.Intensity, "<br>", "<b> Area:</b> ", Sites_eccc$Area.ha, " ha")) %>%
  setView(-123.44799, 48.64919, 9) %>%
  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
  addControl(title, position = "topright")

# Display map
ECCC_map

# Save map
saveWidget(ECCC_map, "output/ECCC_map.html")

```

## Circle Markers
```{r leaflet_map_circle_markers}

# NOT WORKING - Sites not same
#geo_eccc_sites <- cbind(geo_eccc, Sites_eccc)
# try to add , "<br>",  "<b>Composite Index:</b> ", geo_eccc_sites$Composite_Index
# names(geo_eccc_sites)
names(Sites_eccc)

library(dplyr)
library(leaflet)
library(htmlwidgets) # for responsive web map


# Title control
title <- tags$div(HTML('<b>ECC Site Maps</b><br>Malloff & Shackelford. (2024). Feeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. Restoration Futures Lab at the University of Victoria'))
 
# https://rstudio.github.io/leaflet/articles/markers.html 
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("green", "orange"), domain = c("Saanich Peninsula", "Gulf Islands"))

ECCC_map <- leaflet(Sites_eccc) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(
             ~ Lng, ~ Lat, radius = ~ifelse(Subregion == "Saanich Peninsula", 4, 5),
    color = ~pal(Subregion),
    stroke = FALSE, fillOpacity = 0.5,
             popup = paste0("<b>Site:</b> ", Sites_eccc$Site, "<br>", "<b>Subregion:</b> ", Sites_eccc$Subregion, "<br>", "<b>Latitude:</b> ", Sites_eccc$Lat, "<br>", "<b>Longitude:</b> ", Sites_eccc$Lng,  "<br><br>", "<b>Land Manager:</b> ", Sites_eccc$Land.Manager, "<br>", "<b>Main Restoration:</b> ", Sites_eccc$Main.Restoration.Type, "<br>", "<b>Restoration Intensity:</b> ", Sites_eccc$Restoration.Intensity, "<br>", "<b> Area:</b> ", Sites_eccc$Area.ha, " ha")) %>%
  setView(-123.44799, 48.64919, 9) %>%
  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
  addControl(title, position = "topright")

# Display map
ECCC_map

# Save map
saveWidget(ECCC_map, "output/ECCC_map.html")

```

## Circle Markers Sized to Area
```{r leaflet_map_circle_markers_sized_area}

names(Sites_eccc)

library(dplyr)
library(leaflet)
library(htmlwidgets) # for responsive web map


# Title control
# Malloff & Shackelford. (2024). Feeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. Restoration Futures Lab at the University of Victoria
title <- tags$div(HTML('<p style="font-size:12px;font-weight:bold">ECC Sites Map</p>'))
 
# https://rstudio.github.io/leaflet/articles/markers.html 
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("green", "orange"), domain = c("Saanich Peninsula", "Gulf Islands"))

ECCC_map <- leaflet(Sites_eccc) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(
             ~ Lng, ~ Lat, 
# https://stackoverflow.com/questions/40137212/variable-marker-size-feature-in-leaflet-r
              radius = ~geo_eccc$Composite_Index * 3,    
              # radius = ~Sites_eccc$Area.ha,
                  # radius = ~sqrt(Sites_eccc$Area.ha),
 # size smaller is Saanich, else larger if Gulf            
#             radius = ~ifelse(Subregion == "Saanich Peninsula", 4, 5),
    color = ~pal(Subregion),
    stroke = FALSE, fillOpacity = 0.5,
             popup = paste0("<b>Site:</b> ", Sites_eccc$Site, "<br>", "<b>Subregion:</b> ", Sites_eccc$Subregion, "<br>", "<b>Latitude:</b> ", Sites_eccc$Lat, "<br>", "<b>Longitude:</b> ", Sites_eccc$Lng,  "<br><br>", "<b>Land Manager:</b> ", Sites_eccc$Land.Manager, "<br>", "<b>Main Restoration:</b> ", Sites_eccc$Main.Restoration.Type, "<br>", "<b>Restoration Intensity:</b> ", Sites_eccc$Restoration.Intensity, "<br>", "<b> Area:</b> ", Sites_eccc$Area.ha, " ha")) %>%
  setView(-123.44799, 48.64919, 9) %>%
  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
  addControl(title, position = "topright")

# Display map
ECCC_map

# Save map
saveWidget(ECCC_map, "output/ECCC_map.html")

```


## Circle Markers Sized to Composite_Index
geo_eccc_all_site_data.csv
```{r leaflet_map_circle_markers_sized_Composite_Index}

geo_eccc_sites_1 <- read.csv("data/geo_eccc_all_site_data.csv")
names(geo_eccc_sites_1)

library(dplyr)
library(leaflet)
library(htmlwidgets) # for responsive web map


# Title control
# Malloff & Shackelford. (2024). Feeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. Restoration Futures Lab at the University of Victoria
title <- tags$div(HTML('<p style="font-size:12px;font-weight:bold">ECC Sites Map</p>'))
 
# https://rstudio.github.io/leaflet/articles/markers.html 
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("#d7191c", "#2c7bb6"), domain = c("Saanich Peninsula", "Gulf Islands"))

ECCC_map_1 <- leaflet(geo_eccc_sites_1) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(
             ~ Lng, ~ Lat, 
# https://stackoverflow.com/questions/40137212/variable-marker-size-feature-in-leaflet-r
              radius = ~geo_eccc_sites_1$Composite_Index * 5,    
              # radius = ~geo_eccc_sites_1$Area.ha,
                  # radius = ~sqrt(geo_eccc_sites_1$Area.ha),
 # size smaller is Saanich, else larger if Gulf            
#             radius = ~ifelse(Subregion == "Saanich Peninsula", 4, 5),
    fillColor = ~pal(Subregion),
    color = "black",
    weight = 3, # size of circle border
    stroke = TRUE, fillOpacity = 0.5,
             popup = paste0("<b>Site:</b> ", "<b>", "<b>", geo_eccc_sites_1$Site, "</b>", "</b>", "<br><br>", "<b>Subregion:</b> ", geo_eccc_sites_1$Subregion, "<br>",
                        "<b>Latitude:</b> ", geo_eccc_sites_1$Lat, "<br>", "<b>Longitude:</b> ", geo_eccc_sites_1$Lng,  "<br><br>",
                        "<b>Land Manager:</b> ", geo_eccc_sites_1$Land_Manager, "<br>",
                        "<b>Main Restoration:</b> ", geo_eccc_sites_1$Main_Restoration_Type, "<br>",
                        "<b>Restoration Intensity:</b> ", geo_eccc_sites_1$Restoration_Intensity, "<br>",
                        "<b> Area:</b> ", geo_eccc_sites_1$Area_ha, " ha",
                        "<br><br>",
                        "<b>Proportion of native species:</b> ", geo_eccc_sites_1$Proportion_of_native_species, "<br>",
                        "<b>Cultural species richness:</b> ", geo_eccc_sites_1$Cultural_species_richness, "<br>",
                        "<b>Exotic species:</b> ", geo_eccc_sites_1$Exotic_species, "<br>",
                        "<b>Trampling:</b> ", geo_eccc_sites_1$Trampling, "<br>",
                        "<b>Herbivory:</b> ", geo_eccc_sites_1$Herbivory, "<br>",
                        "<b>Composite Index:</b> ", geo_eccc_sites_1$Composite_Index, "<br>")) %>%
  setView(-123.44799, 48.64919, 10) %>%
  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
  addControl(title, position = "topright")

# Display map
ECCC_map_1

# Save map
saveWidget(ECCC_map_1, "output/ECCC_map_CI.html")

```

## Circle Markers Sized to Radius
### Title code
geo_eccc_all_site_data.csv
```{r leaflet_map_circle_markers_sized_radius}

geo_eccc_sites_1 <- read.csv("data/geo_eccc_all_site_data.csv")
names(geo_eccc_sites_1)

library(dplyr)
library(leaflet)
library(htmlwidgets) # for responsive web map


# Title control
# Malloff & Shackelford. (2024). Feeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems. Restoration Futures Lab at the University of Victoria
title <- '<p style="text-align: center; height: 30px; ">
<span style="font-size:10px;font-weight:bold; background-color: rgba(255, 255, 255, 0.9;");>ECCC GOE 2023 Monitoring Sites</span><br>
<span style="font-size:8px;font-style:italic; background-color: rgba(255, 255, 255, 0.9;">(Malloff & Shackelford, 2024)</span></p>'
 
# https://rstudio.github.io/leaflet/articles/markers.html 
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("#d7191c", "#2c7bb6"), domain = c("Saanich Peninsula", "Gulf Islands"))

ECCC_map_2 <- leaflet(geo_eccc_sites_1) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(
             ~ Lng, ~ Lat, 
# https://stackoverflow.com/questions/40137212/variable-marker-size-feature-in-leaflet-r
#              radius = ~geo_eccc_sites_1$Composite_Index * 5,    
              # radius = ~geo_eccc_sites_1$Area.ha,
                  # radius = ~sqrt(geo_eccc_sites_1$Area.ha),
 # size smaller is Saanich, else larger if Gulf            
             radius = ~ifelse(Subregion == "Saanich Peninsula", 8, 8),
    fillColor = ~pal(Subregion),
    color = "black",
    weight = 3, # size of circle border
    stroke = TRUE, fillOpacity = 0.5,
             popup = paste0("<b>ECCC GOE 2023 Monitoring Data</b>", "<br>", "<i>(Malloff & Shackelford, 2024)</i>", "<br><br>",
               "<b>Site:</b> ", "<b>", geo_eccc_sites_1$Site, "</b>", "</b>", "<br><br>", "<b>Subregion:</b> ", geo_eccc_sites_1$Subregion, "<br>",
                        "<b>Latitude:</b> ", geo_eccc_sites_1$Lat, "<br>", "<b>Longitude:</b> ", geo_eccc_sites_1$Lng,  "<br><br>",
                        "<b>Land Manager:</b> ", geo_eccc_sites_1$Land_Manager, "<br>",
                        "<b>Main Restoration:</b> ", geo_eccc_sites_1$Main_Restoration_Type, "<br>",
                        "<b>Restoration Intensity:</b> ", geo_eccc_sites_1$Restoration_Intensity, "<br>",
                        "<b> Area:</b> ", geo_eccc_sites_1$Area_ha, " ha",
                        "<br><br>",
                        "<b>Proportion of native species:</b> ", geo_eccc_sites_1$Proportion_of_native_species, "<br>",
                        "<b>Cultural species richness:</b> ", geo_eccc_sites_1$Cultural_species_richness, "<br>",
                        "<b>Exotic species:</b> ", geo_eccc_sites_1$Exotic_species, "<br>",
                        "<b>Trampling:</b> ", geo_eccc_sites_1$Trampling, "<br>",
                        "<b>Herbivory:</b> ", geo_eccc_sites_1$Herbivory, "<br>",
                        "<b>Composite Index:</b> ", geo_eccc_sites_1$Composite_Index, "<br>")) %>%
  setView(-123.44799, 48.64919, 10) %>%
  # add controls
  addMiniMap(width = 150, height = 150, zoomLevelOffset = -4) %>%
  addControl(title, position = "topright")

# Display map
ECCC_map_2

# Save map
saveWidget(ECCC_map_2, "output/ECCC_map_radius-1.html")

```

---

# Shiny 
## Shiny Gadget
### lmGadget all sites
- code works but can't knit html file unless
adding "echo=TRUE, eval = FALSE"
```{r shiny_lmgadget, echo=TRUE, eval = FALSE}
# colnames(geo_eccc_1)

library(shiny)
library(miniUI)
library(ggplot2)


lmGadget <- function(data, xvar, yvar) {

  ui <- miniPage(
    gadgetTitleBar("Interactive lm"),
    miniContentPanel(
      fillRow(flex = c(NA, 1),
              fillCol(width = "100px",
                      selectInput("degree", "Polynomial degree", c(1, 2, 3, 4))
              ),
              plotOutput("plot1",
                         height = "100%",
                         click = "plot1_click",
                         brush = brushOpts(
                           id = "plot1_brush"
                         )
              )
      )
    ),
    miniButtonBlock(
      actionButton("exclude_toggle", "Toggle points"),
      actionButton("exclude_reset", "Reset")
    )
  )

  server <- function(input, output) {
    # For storing which rows have been excluded
    vals <- reactiveValues(
      keeprows = rep(TRUE, nrow(data))
    )

    output$plot1 <- renderPlot({
      req(input$degree)
      formula <- as.formula(paste0("y ~ poly(x, degree = ", input$degree, ")"))

      # Plot the kept and excluded points as two separate data sets
      keep    <- data[ vals$keeprows, , drop = FALSE]
      exclude <- data[!vals$keeprows, , drop = FALSE]

      ggplot(keep, aes_string(xvar, yvar)) + geom_point() +
        geom_smooth(method = lm, formula = formula, fullrange = TRUE, color = "gray50") +
        geom_point(data = exclude, fill = NA, color = "black", alpha = 0.25) +
        coord_cartesian(xlim = range(data[[xvar]]), ylim = range(data[[yvar]])) +
        theme_bw(base_size = 14)
    })

    # Toggle points that are clicked
    observeEvent(input$plot1_click, {
      res <- nearPoints(data, input$plot1_click, allRows = TRUE)

      vals$keeprows <- xor(vals$keeprows, res$selected_)
    })

    # Toggle points that are brushed, when button is clicked
    observeEvent(input$exclude_toggle, {
      res <- brushedPoints(data, input$plot1_brush, allRows = TRUE)

      vals$keeprows <- xor(vals$keeprows, res$selected_)
    })

    # Reset all points
    observeEvent(input$exclude_reset, {
      vals$keeprows <- rep(TRUE, nrow(data))
    })

    # Handle the Done button being pressed.
    observeEvent(input$done, {
      # Replace x and y in the formula with the values in xvar and yvar
      formula <- as.formula(paste0(yvar, " ~ poly(", xvar, ", degree = ", input$degree, ")"))
      keep_data <- data[vals$keeprows, , drop = FALSE]

      # Return the kept points.
      stopApp(
        list(
          data = keep_data,
          model = lm(formula, keep_data)
        )
      )
    })

  }

    runGadget(ui, server, viewer = dialogViewer("lmGadget")) # separate viewer window
}

lmGadget(geo_eccc_1, "Percentage_ns", "Composite_Index")
lmGadget(iris, "Sepal.Length", "Sepal.Width")
```



## Inline Shiny app 
### Reactive choose 
```{r, Shiny_app_inline, echo=FALSE}
library(shiny)
geo_eccc <- read.csv("data/SiteData-TEST-GOE-Monitor.csv", header = TRUE, sep = ",", stringsAsFactors=FALSE)
colnames(geo_eccc)
str(geo_eccc)

shinyApp(

# https://stackoverflow.com/questions/44569739/r-shiny-how-to-make-selectinput-choices-reactive-to-each-other-while-subsetting




ui <-shinyUI(fluidPage(
  selectInput("Box1","Choose a Subregion", choices = c("All",unique(geo_eccc$Subregion))),
  selectInput("Box2","Choose an Site", choices = c("All",unique(geo_eccc$Site))),
  tableOutput("table1")
)), # include comma after ui


server <- shinyServer(function(input,output, session){

  data1 <- reactive({
    if(input$Box1 == "All"){
      geo_eccc
    }else{
      geo_eccc[which(geo_eccc$Subregion == input$Box1),]
    }
  })

  data2 <- reactive({
    if (input$Box2 == "All"){
      geo_eccc
    }else{
      geo_eccc[which(geo_eccc$Site == input$Box2),]
    }
  })

  observe({

    if(input$Box1 != "All"){
      updateSelectInput(session,"Box2","Choose an Site", choices = c("All",unique(data1()$Site)))
    }

    else if(input$Box2 != 'All'){
      updateSelectInput(session,"Box1","Choose a Subregion", choices = c('All',unique(data2()$Subregion)))
    }

    else if (input$Box1 == "All" & input$Box2 == "All"){
      updateSelectInput(session,"Box2","Choose an Site", choices = c('All',unique(geo_eccc$Site)))
      updateSelectInput(session,"Box1","Choose a Subregion", choices = c('All',unique(geo_eccc$Subregion)))
    }
  })


  data3 <- reactive({
    if(input$Box2 == "All"){
      data1()
    }else if (input$Box1 == "All"){
      data2()
    }else if (input$Box2 == "All" & input$Box1 == "All"){
      geo_eccc
    }
    else{
      geo_eccc[which(geo_eccc$Site== input$Box2 & geo_eccc$Subregion == input$Box1),]
    }
  })

  output$table1 <- renderTable({
    data3()
  })

})

)
```


### Leaflet Reactive Site Details
`app-leaflet-reactive-AllSiteDetails-WORKING`

```{r, Shiny_app_inline_leaflet, echo=FALSE}
## https://stackoverflow.com/questions/46979328/how-to-make-shiny-leaflet-map-react-to-change-in-input-value-r
# Map is filtering Subregion Sites, but drop down list isn't
# Modelled on `app-test-leaflet-reactive.R`
# Works 2025-08-18

library(shiny)
library(bslib)
library(leaflet)
library(dplyr)

############################################################
## Data
# note Restoration_Intensity column values all had a space in front of value - wasn't sorting properly ... changed
geo_eccc_sites <- read.csv("data/geo_eccc_all_site_data.csv", header = TRUE, sep = ",")
names(geo_eccc_sites)
str(geo_eccc_sites)

shinyApp(
############################################################
## UI
ui <- fluidPage(
  titlePanel("ECCC GOE Monitoring Sites Interactive Map"),
  tags$div(HTML('<p style="font-size:12px;"><b>Source:</b> Malloff & Shackelford. (2024).
  <i>Feeling the Pulse: Monitoring methods and initial outcomes in oak meadow ecosystems.</i>
                Restoration Futures Lab at the University of Victoria.</p>')),
  tags$div(HTML('<p style="font-size:12px;"><b>Note:</b> Marker size is based on Composite Index.</p>')),

    # Dropdown for Subregion
    selectInput(
      "Subregion_filter",
      "Select Subregion:",
      choices = c("All" = "all", unique(geo_eccc_sites$Subregion)),
      selected = "all"
    ),
  tags$div(HTML('<p style="font-size:12px;">Note: If selecting by Subregion,
                ensure Sites are All selected.</p>')),

    # Dropdown for Site
    selectInput(
      "Site_filter",
      "Select Site:",
      choices = c("All" = "all", unique(geo_eccc_sites$Site)),
      selected = "all"
  ),

    leafletOutput("map", height = "600px")
),

############################################################
## Server
server <- function(input, output, session) {

  pal <- colorFactor(c("green", "orange"), domain = c("Saanich Peninsula", "Gulf Islands"))
  # Reactive expression to filter data based on dropdown selections
  filtered_data <- reactive({
    data <- geo_eccc_sites

    # Filter by Subregion if not "all"
    if (input$Subregion_filter != "all") {
      data <- data %>% filter(Subregion == input$Subregion_filter)
    }

    # Filter by Site if not "all"
    if (input$Site_filter != "all") {
      data <- data %>% filter(Site == input$Site_filter)
    }

    return(data)
  })

  # Create the base map
  output$map <- renderLeaflet({
    leaflet() %>%
      addProviderTiles("Esri.WorldImagery") %>%
      setView(-123.44799, 48.64919, 9) # Center of US
  })

  # Observer to update markers when filtered data changes
  observe({
    data <- filtered_data()

    leafletProxy("map") %>%
      clearMarkers() %>%  # Remove existing markers
      addCircleMarkers(
        color = ~pal(Subregion),
        # size by multiple of Composite_Index
        radius = ~Composite_Index * 5,
        stroke = FALSE, fillOpacity = 0.5,
        data = data,
        lng = ~Lng,
        lat = ~Lat,
        popup = ~paste0("<b>Site:</b> ", "<b>", Site, "</b>", "<br><br>", "<b>Subregion:</b> ", Subregion, "<br>",
                        "<b>Latitude:</b> ", Lat, "<br>", "<b>Longitude:</b> ", Lng,  "<br><br>",
                        "<b>Land Manager:</b> ", Land_Manager, "<br>",
                        "<b>Main Restoration:</b> ", Main_Restoration_Type, "<br>",
                        "<b>Restoration Intensity:</b> ", Restoration_Intensity, "<br>",
                        "<b> Area:</b> ", Area_ha, " ha",
                        "<br><br>",
                        "<b>Proportion of native species:</b> ", Proportion_of_native_species, "<br>",
                        "<b>Cultural species richness:</b> ", Cultural_species_richness, "<br>",
                        "<b>Exotic species:</b> ", Exotic_species, "<br>",
                        "<b>Trampling:</b> ", Trampling, "<br>",
                        "<b>Herbivory:</b> ", Herbivory, "<br>",
                        "<b>Composite Index:</b> ", Composite_Index, "<br>"
                        )
      )
  })
}
)
#shinyApp(ui = ui, server = server)

```


### Datatable search plot 
`app-datatable-search-plot-update-ECCC-TEST`
https://stackoverflow.com/questions/52880657/in-an-shiny-app-i-want-a-plot-to-update-based-on-the-search-results-in-a-datat 
```{r, Shiny_app_inline_datatable, echo=FALSE}

library(shiny)
library(DT)

geo_eccc <- read.csv("data/geo_eccc_all_site_data.csv", header = TRUE, sep = ",", stringsAsFactors=FALSE)

shinyApp(
ui <- basicPage(
  h2("ECCC GOE Monitoring Data"),
  DT::dataTableOutput("mytable"),
  downloadButton("downloadData", label = "Download filtered table data"),
  tags$div(HTML("<br>")),
  verbatimTextOutput("test"),
  plotOutput('plot1')

),

server <- function(input, output) {

  mc <- geo_eccc # could be reactive in real world case

  output$mytable = DT::renderDataTable({
    datatable(mc, filter = 'top')
  })

  filtered_table <- reactive({
    req(input$mytable_rows_all)
    mc[input$mytable_rows_all, ]
  })

  output$plot1 <- renderPlot({
    plot(filtered_table()$Proportion_of_native_species, filtered_table()$Exotic_species, col = "green", lwd = 10)
    #    plot(filtered_table()$Proportion_of_native_species, filtered_table()$Exotic_species, col = "red", lwd = 10)
  })

  output$test <- renderPrint({
    filtered_table()
  })

  output$downloadData <- downloadHandler(
    filename = "myFile.csv",
    content = function(file){
      write.csv(filtered_table(), file)
    }
  )

}
)
# shinyApp(ui, server)

```


  ### Plot Select Axis Variables
`app4-SelectAxisVariables-WORKS.R`
- this shiny app won't knit - looses connection ???
```{r, Shiny_app_inline_, echo=FALSE}
#
# https://monashbioinformaticsplatform.github.io/R-ShinyIntro-MBP/tutorial.html
# https://github.com/MonashDataFluency/R-ShinyIntro/blob/master/ShinyApps/demo/demo_interactive.R

# https://monashbioinformaticsplatform.github.io/R-ShinyIntro-MBP/tutorial.html

library(shiny)
library(shinyscreenshot)
library(utils)
library(ggplot2)

# shinyscreenshot::runExample()
shinyApp(
ui <- fluidPage(
  titlePanel("Plotting ECCC GOE Monitoring Variables"),

  sidebarLayout(
    sidebarPanel(
      # set div for screenshot
        div(
          id = "input_panel",
          tags$b("Input sidebar panel selections"),
          selectInput(inputId = "x_select", label = "Select x-axis",
                      choices = list("Proportion_of_native_species",
                                     "Cultural_species_richness",
                                     "Exotic_species",
                                     "Trampling",
                                     "Herbivory",
                                     "Composite_Index"),
                      selected = "Exotic_species"),
          selectInput(inputId = "y_select", label = "Select y-axis",
                      choices = list("Proportion of native species" = "Proportion_of_native_species",
                                     "Cultural species richness" = "Cultural_species_richness",
                                     "Exotic species" = "Exotic_species",
                                     "Trampling 1" = "Trampling",
                                     "Herbivory 1" = "Herbivory",
                                     "Composite Index" = "Composite_Index"),
                      selected = "Proportion_of_native_species"),
        ),
        tags$div(HTML("<hr style='height:2px;border-width:0;color:gray;background-color:gray'>")),
## Screenshot buttons
                 tags$div(HTML("<b>Take Screenshots<br></b>")),
                 tags$div(HTML("<p style='font-size:10px;font-style:italic;'>(Note: Screenshots are saved to same location as Shiny app)</p>")),
                 screenshotButton(label = "Entire page", filename = "ECCC-Screenshot-entirePage", download = FALSE, server_dir="."),
                 screenshotButton(label = "Input panel", id = "input_panel", filename = "ECCC-Screenshot-inputPanel", download = FALSE, server_dir="."),
                 screenshotButton(label = "Plot 1", id = "my_plot", filename = "ECCC-Screenshot-plot1", download = FALSE, server_dir="."),
                 screenshotButton(label = "Plot 2", id = "my_plot1", filename = "ECCC-Screenshot-plot2", download = FALSE, server_dir=".")
    ),
    mainPanel(tags$b("Main panel Input Selections"),
              tags$div(HTML("<br>")),
              strong("X axis: "),
              textOutput(outputId = "check_x_select"),
              tags$div(HTML("<br>")),
              strong("Y axis: "),
              textOutput(outputId = "check_y_select"),
              #tags$div(HTML('<br><br>')),
              tags$div(HTML("<hr style='height:2px;border-width:0;color:gray;background-color:gray'>")),
              tags$div(HTML("<b>", "Plot1: Resulting ggplot by Site", "</b>")),
              plotOutput(outputId = "my_plot", width = 600, height = 600),
              tags$div(HTML("<hr style='height:2px;border-width:0;color:gray;background-color:gray'>")),
              tags$div(HTML("<b>", "Plot2: Resulting ggplot by Subregion", "</b>")),
              plotOutput(outputId = "my_plot1", width = 600, height = 600)
    )
  )
),

server <- function(input, output) {

  output$check_x_select <- renderText({
    print(input$x_select)
  })

  output$check_y_select <- renderText({
    print(input$y_select)
  })

  output$my_plot <- renderPlot({

    #Read in the sample csv
    dat <- read.csv("data/TEST-GOE-Monitor.csv")

    #Pull out a column for the x and y axis each
    x_axis <- dat[, input$x_select]
    y_axis <- dat[, input$y_select]

    #Create a basic dot plot
    #plot(x = x_axis, y_axis)

    # ggplot
    ggplot() +
           aes(x = x_axis, y_axis, color = dat$Site) +
      geom_point(shape = 18, size = 4) +
      theme_minimal() + #get rid of grey background and tick marks
      theme(legend.position="bottom") + #remove legend
      theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
      labs(title='Plotting Site Variables by Select Input',
           subtitle='ECCC GEO Monitoring Data',
           caption = "Chart by Wendy Anthony \n 2025-08-18",
            x = input$x_select, y = input$y_select,
           # legend title by variable used to colour
           color = "Site")
  })

  output$check_y_select <- renderText({
    print(input$y_select)
  })

  output$my_plot1 <- renderPlot({

    #Read in the sample csv
    dat <- read.csv("data/TEST-GOE-Monitor.csv")

    #Pull out a column for the x and y axis each
    x_axis <- dat[, input$x_select]
    y_axis <- dat[, input$y_select]

    #Create a basic dot plot
    #plot(x = x_axis, y_axis)

    # ggplot
    ggplot() +
      aes(x = x_axis, y_axis, color = dat$Subregion) +
      geom_point(shape = 18, size = 4) +
      theme_minimal() + #get rid of grey background and tick marks
      theme(legend.position="bottom") + #remove legend
      theme(axis.text.x = element_text(angle = 25, vjust = 1, hjust=1, size = 7)) +
      labs(title='Plotting Subregion Variables by Select Input',
           subtitle='ECCC GEO Monitoring Data',
           caption = "Chart by Wendy Anthony \n 2025-08-18",
           x = input$x_select, y = input$y_select,
           # legend title by variable used to colour
           color = "Subregion")
  })

}
)
# shinyApp(ui = ui, server = server)

```


---

# TO DO
- leaflet map jumps when clicking popup (in RStudio, not in html file)

## DONE
### to combine dataframes csv files to include site info
1. make sure they each have same length number of unique rows
2. arrange alphabetic order
3. cbind
4. remove duplicate columns

# Session info
```{r session-info}
# to document specific packages used to run script
sessionInfo()
```








